### 2020-11-04 HCM.SCD.CEanalysis/README.Rmd
add to README

## Bayesian Models for CE Analysis of a Prognostic Model for Sudden Cardiac Death (SCD) in Hypertrophic Cardiomyopathy (HCM)

### Background

Hypertrophic cardiomyopathy(HCM) is a leading cause of sudden cardiac death (SCD) in young adults. Current risk algorithms provide only a crude estimate of risk and fail to account for the different effect size of individual risk factors.

### Aim

The aim of this study was to perform a cost-effectiveness analysis for the new SCD risk prediction model. See for more information: [C O'Mahony (2014) A novel clinical risk prediction model for sudden cardiac death in hypertrophic cardiomyopathy (HCM risk-SCD). Eur Heart;35(30):2010-20.](https://pubmed.ncbi.nlm.nih.gov/24126876/)

### Overview

We develop a predictive model for assessing risk of SCD among patients with HCM.

-   Using individual level data generate posterior distribution samples of probabilities, costs and health values using WinBUGS
-   Simulate with a Markov model using the posterior samples as inputs
-   Compare outcomes for risk groups in terms of cost-effectiveness

The model structure consists of a decision tree component and a Markov model component. In effect this first model is represented by different cohorts of the data in the fitting process.

<img src="images/model_diagram.png" alt="drawing" width="500"/>

### Data

Individual HCM patient level data.

| Variable     | Description                                                                  |
---------------|-------------------------------------------------------------------------------
| set          | "Imputed dataset"                                                            |
| centre       | "Centre"                                                                     |
| id           | "ID"                                                                         |
| d            | "Censoring Indicator (1=SCD/ICD shock)"                                      |
| time         | "Time (years)"                                                               |
| age          | "Age (years)"                                                                |
| mwt          | "Maximal wall thickness (mm)"                                                |
| mwt2         | "mwt^2"                                                                      |
| la           | "Left atrial size (mm)"                                                      |
| maxlvotg     | "Maximal LVOT gradient (mmHg)"                                               |
| fhxscd       | "Family histroy of SCD"                                                      |
| nsvt         | "Non Sustained Ventricular Tachycardia"                                      |
| syncope      | "Syncope (blackout)"                                                         |
| na           | "Nelson-Aalen cumulative hazard"                                             |
| year         | "Year of entry in study"                                                     |
| male         | "Male"                                                                       |
| nyha3_4      | "NYHA class 3/4"                                                             |
| mye          | "Myectomy"                                                                   |
| mye_t0       | "Myectomy at baseline"                                                       |
| myedate      | "Myectomy date"                                                              |
| asa          | "Alcohol septal ablation"                                                    |
| asa_t0       | "ASA at baseline"                                                            |
| asadate      | "Date of ASA"                                                                |
| icd          | "Implantable Cardioverter defibrillator"                                     |
| icd_t0       | "ICD at baseline"                                                            |
| amio         | "Amiodarone"                                                                 |
| perm_persaf  | "Permanent or Persistent Atrial Fibrillation"                                |
| mwt30        | "MWT = or > 30mm"                                                            |
| lvedd        | "LV end diastolic dimension (mm)"                                            |
| fs           | "Fractional Shortening (%)"                                                  |
| lvoto30      | "LVOT gradient = or >30mmHg"                                                 |
| lvoto50      | "LVOT gradient = or >50mmHg"                                                 |
| t0           | "Date of baseline evaluation"                                                |
| tmax         | "Date of censoring or endpoint"                                              |
| yearend      | "Year exiting study"                                                         |
| fu           | "Follow-up period (years)"                                                   |
| scd          | "Sudden Cardiac Death"                                                       |
| scdtype      | "SCD type: 1= appropriate ICD shock, 2=survived cardiac arrest, 3=actual SCD"|
| cvs_all      | "All cardiovascular deaths combined"                                         |
| ccf_tx       | "Heart failure deaths and transplants"                                       |
| ccf          | "Heart failure deaths only"                                                  |
| tx           | "Transplants only"                                                           |
| cva          | "Stroke deaths"                                                              |
| cvs_other    | "Other cardiovascular deaths"                                                |
| non_cvs      | "Non cardiovascular deaths"                                                  |
| unknowndeath | "Unknown Deaths (grouped with non_cvs)"                                      |
| anyrip_tx    | "All deaths and transplants"                                                 |
| centre2      | "Centre"                                                                     |
| centre_cat1  | "centre2==Athens"                                                            |
| centre_cat2  | "centre2==Bologna"                                                           |
| centre_cat3  | "centre2==Coruna"                                                            |
| centre_cat4  | "centre2==London"                                                            |
| centre_cat5  | "centre2==Murcia"                                                            
| centre_cat6  | "centre2==Naples"

### 2020-11-04 HCM.SCD.CEanalysis/scripts/prep_study_data.R

@@ -24,6 +24,7 @@ mydata <- read_dta("raw data/hcmdata.dta")
 data_set1 <- mydata %>% filter(set == 1)
 
 save(data_set1, file = "data/data_set1.RData")
+
 CYCLE <- 1
 
 # non-deterministic decision

### 2020-11-05 HCM.SCD.CEanalysis/R/annual_trans_counts.R
add censoring to obs_aggr_trans_mat()

@@ -3,6 +3,10 @@
 #'
 #' Aggregate individual data to annual totals.
 #'
+#' @param data
+#' @param cycle_length
+#' @export
+#'
 annual_trans_counts <- function(data,
                                 cycle_length = 1) {
 
@@ -11,17 +15,20 @@ annual_trans_counts <- function(data,
   data %>%
     # create combined non-scd field
     mutate(
-      non_scd = as.numeric((cvs_all & !scd) | non_cvs)) %>%
-    select(time, non_scd, scd) %>%
+      non_scd = as.numeric((cvs_all & !scd) | non_cvs),
+      cens = as.numeric(!non_scd & !scd)) %>%
+    select(time, non_scd, scd, cens) %>%
     mutate(yr_grp = cut(time,
-                        breaks = seq(0, 35, by = cycle_length),
+                        breaks = seq(0, 40, by = cycle_length),
                         right = FALSE)) %>%
     group_by(yr_grp) %>%
     summarise(scd_count = sum(scd),
-              non_scd_count = sum(non_scd)) %>%
+              non_scd_count = sum(non_scd),
+              cens_count = sum(cens)) %>%
     mutate(cum_scd = cumsum(scd_count),
            cum_non_scd = cumsum(non_scd_count),
-           healthy = n_pop - cum_scd - cum_non_scd,
+           cum_cens = cumsum(cens_count),
+           healthy = n_pop - cum_cens - cum_scd - cum_non_scd,
            at_risk = lag(healthy, default = n_pop))
 }
 

### 2020-11-05 HCM.SCD.CEanalysis/R/obs_aggr_trans_mat.R

@@ -1,34 +1,41 @@
 
-#' obs_aggr_tran_mat
+#' obs_aggr_trans_mat
 #'
 #' Create aggregate observed transition counts
-#' see Table 5.2 chapter 5 in BCEA book
+#' see p.185 Table 5.2 chapter 5 in BCEA book
 #'
 #' @param data new state arrivals at time step
 #'             e.g. data = data_rule$`FALSE`
+#' @param max_year
+#' @export
 #'
-obs_aggr_tran_mat <- function(data,
-                              max_year = 5) {
+obs_aggr_trans_mat <- function(data,
+                               max_year = 5) {
 
   trans_mat <- list()
   state_names <- c("healthy", "scd_count", "non_scd_count")
 
+  # initialise counts
   trans_mat[[1]] <-
     rbind(c(0,0,0),
           c(0,0,0),
           c(0,0,0)) %>%
-    `colnames<-`(state_names)
+    `colnames<-`(state_names) %>%
+    `rownames<-`(state_names)
 
   for (i in seq_len(max_year)) {
 
     trans_mat[[i + 1]] <-
+      # previous transitions
       trans_mat[[i]] +
+      # new transitions
       rbind(data[i, state_names],
             c(0,0,0),
             c(0,0,0)) +
-      rbind(c(0,0,0),
-            c(0, trans_mat[[i]][1, "scd_count"], 0),
-            c(0,0, trans_mat[[i]][1, "non_scd_count"]))
+      # absorbing states
+      diag(c(0,
+             trans_mat[[i]]["healthy", "scd_count"],
+             trans_mat[[i]]["healthy", "non_scd_count"]))
   }
 
   names(trans_mat) <- c(0:max_year)

### 2020-11-05 HCM.SCD.CEanalysis/scripts/prep_study_data.R

@@ -30,13 +30,15 @@ CYCLE <- 1
 # non-deterministic decision
 fuzzy_risk_noise <- rnorm(nrow(data_set1), 0, 0.001)
 
+
 # cox risk score > 6% ----
 
 data_set1$risk_over_6 <-
   as.factor((data_set1$risk_5_years + fuzzy_risk_noise) > 0.06)
 
 data_risk6 <-
-  group_split(data_set1, risk_over_6) %>%
+  data_set1 %>%
+  group_split(risk_over_6) %>%
   setNames(levels(data_set1$risk_over_6)) %>%
   map(.f = annual_trans_counts,
       cycle_length = CYCLE) %>%

### 2020-11-06 HCM.SCD.CEanalysis/BUGS/script.R
simplify transition count matrix function

@@ -2,6 +2,10 @@
 # run BUGS model script
 # cf BCEA book chapter 5
 #
+# estimate transition probabilities
+# from discrete time counts
+# using multinomial likelihood
+
 
 
 library(R2jags)

### 2020-11-06 HCM.SCD.CEanalysis/R/obs_aggr_trans_mat.R

@@ -25,17 +25,20 @@ obs_aggr_trans_mat <- function(data,
 
   for (i in seq_len(max_year)) {
 
-    trans_mat[[i + 1]] <-
-      # previous transitions
-      trans_mat[[i]] +
-      # new transitions
-      rbind(data[i, state_names],
-            c(0,0,0),
-            c(0,0,0)) +
-      # absorbing states
-      diag(c(0,
-             trans_mat[[i]]["healthy", "scd_count"],
-             trans_mat[[i]]["healthy", "non_scd_count"]))
+    trans_mat[[i + 1]] <- trans_mat[[i]]
+
+    # new transitions
+      trans_mat[[i + 1]]["healthy", state_names] <-
+        trans_mat[[i + 1]]["healthy", state_names] +
+        unlist(data[i, state_names])
+
+      trans_mat[[i + 1]]["scd_count", "scd_count"] <-
+      trans_mat[[i + 1]]["scd_count", "scd_count"] +
+        trans_mat[[i]]["healthy", "scd_count"]
+
+      trans_mat[[i + 1]]["non_scd_count", "non_scd_count"] <-
+      trans_mat[[i + 1]]["non_scd_count", "non_scd_count"] +
+        trans_mat[[i]]["healthy", "non_scd_count"]
   }
 
   names(trans_mat) <- c(0:max_year)

### 2020-11-06 HCM.SCD.CEanalysis/scripts/prep_study_data.R

@@ -18,23 +18,28 @@ library(haven)
 library(dplyr)
 library(purrr)
 
-mydata <- read_dta("raw data/hcmdata.dta")
+## STATA data
+# mydata <- read_dta("raw data/hcmdata.dta")
+# # single imputed data set
+# data_set1 <- mydata %>% filter(set == 1)
+# save(data_set1, file = "data/data_set1.RData")
 
-# single imputed data set
-data_set1 <- mydata %>% filter(set == 1)
+data("data_set1")
 
-save(data_set1, file = "data/data_set1.RData")
+CYCLE <- 1 #year
 
-CYCLE <- 1
+# non-deterministic decision rule
+FUZZY_RISK <- TRUE
 
-# non-deterministic decision
-fuzzy_risk_noise <- rnorm(nrow(data_set1), 0, 0.001)
-
+fuzzy_noise <-
+  if (FUZZY_RISK) {
+    rnorm(nrow(data_set1), 0, 0.001)
+  } else {0}
 
 # cox risk score > 6% ----
 
 data_set1$risk_over_6 <-
-  as.factor((data_set1$risk_5_years + fuzzy_risk_noise) > 0.06)
+  as.factor((data_set1$risk_5_years + fuzzy_noise) > 0.06)
 
 data_risk6 <-
   data_set1 %>%
@@ -46,22 +51,23 @@ data_risk6 <-
 
 data_risk6
 
-save(data_risk6, file = "data/data_risk6.RData")
+save(data_risk6, file = "data/trans_counts_risk6.RData")
 
 # cox risk score > 4% ----
 
 data_set1$risk_over_4 <-
-  as.factor((data_set1$risk_5_years + fuzzy_risk_noise) > 0.04)
+  as.factor((data_set1$risk_5_years + fuzzy_noise) > 0.04)
 
 data_risk4 <-
   group_split(data_set1, risk_over_4) %>%
   setNames(levels(data_set1$risk_over_4)) %>%
   map(.f = annual_trans_counts,
-      cycle_length = CYCLE)
+      cycle_length = CYCLE) %>%
+  map(.f = obs_aggr_trans_mat)
 
 data_risk4
 
-save(data_risk4, file = "data/data_risk4.RData")
+save(data_risk4, file = "data/trans_counts_risk4.RData")
 
 # status-quo as risk factor rule ----
 #
@@ -71,21 +77,24 @@ save(data_risk4, file = "data/data_risk4.RData")
 # syncope: unexplained syncope
[…]

### 2020-11-06 HCM.SCD.CEanalysis/BUGS/script.R
first complete fitting and ce model run. want to change to have ICD and non-ICD in same sim model do can then compare with different strategies/thresholds/rules

@@ -7,68 +7,66 @@
 # using multinomial likelihood
 
 
-
 library(R2jags)
 
-load("data/data_obs.RData")
+# select:
+data("trans_counts_obs")
+data("trans_counts_risk6")
 
 
 n_S <- 3  # number of states
 J <- 5    # number of time points
 
-r.0 <-
-  data_obs$`0` %>%
-  filter(yr_grp == "[0,5)") %>%
-  select(scd_count, non_scd_count, healthy)
+r.0 <- data_obs$`0`[, c("healthy", "scd_count", "non_scd_count")]
+r.1 <- data_obs$`1`[, c("healthy", "scd_count", "non_scd_count")]
 
-r.1 <-
-  data_obs$`1` %>%
-  filter(yr_grp == "[0,5)") %>%
-  select(scd_count, non_scd_count, healthy)
-
-n.0 <- rowSums(r.0)
-n.1 <- rowSums(r.1)
+n.0 <- unname(rowSums(r.0))
+n.1 <- unname(rowSums(r.1))
 
 scale <- 1                            # level of informativeness for
 alpha.0 <- alpha.1 <- rep(scale, n_S) # the Dirichlet prior
 
 dataJags <-
-  list("n.0", "n.1",
-       "r.0", "r.1",
-       "alpha.0", "alpha.1",
-       "n_S")
+  list(n.0 = n.0,
+       n.1 = n.1,
+       r.0 = r.0,
+       r.1 = r.1,
+       alpha.0 = alpha.0,
+       alpha.1 = alpha.1,
+       n_S = n_S)
 
-filein <- "model.txt"
+filein <- "BUGS/model.txt"
 params <- c("lambda.0", "lambda.1")
 
+#
 inits <- function() {
 
-  temp.0 <- matrix(
-    rgamma(4*n_S, scale, 1) , 4, n_S)
+  temp.0 <- rgamma(n_S, scale, 1)
+  sum.temp.0 <- sum(temp.0)
+  p.0 <- temp.0/sum.temp.0
 
-  sum.temp.0 <- apply(temp.0, 1, sum)
+  temp.1 <- rgamma(n_S, scale, 1)
+  sum.temp.1 <- sum(temp.1)
+  p.1 <- temp.1/sum.temp.1
 
-  mat.0 <- temp.0/sum.temp.0
-
-  temp.1 <-
-    matrix(rgamma(4*n_S, scale, 1), 4, n_S)
-
-  sum.temp.1 <- apply(temp.1, 1, sum)
-
[…]

### 2020-11-06 HCM.SCD.CEanalysis/R/ce_sim.R

@@ -28,7 +28,7 @@ ce_sim <- function(pop,
   n_sim <- dim(pop[[1]])[3]
   n_interv <- length(pop)
 
-  if (is.na(pdecr)) {
+  if (any(is.na(pdecr))) {
     pdecr <- map(1:n_interv, ~rep(0, S))}
 
   # initialise empty output matrices

### 2020-11-06 HCM.SCD.CEanalysis/scripts/main_ce_analysis.R

@@ -7,64 +7,74 @@
 library(dplyr)
 library(purrr)
 library(BCEA)
-library(reshape2)
+library(res_newhape2)
 library(HCM.SCD.CEanalysis)
 
+data(mm1)
+R2WinBUGS::attach.bugs(mm1$BUGSoutput)
 
 # 1. alive with HCM
 # 2. SCD
 # 3. all-cause mortality
 
 n_init <- c(1000, 0, 0)
-n_sim <- 2
+n_sim <- nrow(lambda.0)
 S <- length(n_init)
 J <- 12  # max time
 
 # utilities and costs
-e_unit <- list(c(0.637, 0, 0),
-               c(0.637, 0, 0))
+# for each intervention
 
-c_unit <- list(c(4792, 0, 0),
-               c(4812, 0, 0))
+# 0.637: managed with an ICD (Noyes 2007)
+
+e_unit <- list(low_risk = c(1, 0, 0),
+               ICD = c(0.637, 0.5, 0))
+
+# cost of SCD risk algorithm £20
+# cost of SCD risk management for only day case procedures_new £4792 (Waight 2019)
+# cost of non-fatal HCM related events £22,880 (UK Stroke Association)
+
+c_unit <- list(low_risk = c(0, 0, 0),
+               ICD = c(0, 22880, 0))
 
 # linear proportion decrease each year in state
-pdecr <- list(c(0, 0, 0),
-              c(0, 0, 0))
+pdecr <- list(low_risk = c(0, 0, 0),
+              ICD = c(0, 0, 0))
 
-probs <- list(array(c(0.90, 0.075, 0.025,
-                      0, 1, 0,
-                      0, 0, 1),               # status-quo
-                    dim = c(S, S, 1, n_sim)),
-              array(c(0.90, 0.075, 0.025,
-                      0, 1, 0,
-                      0, 0, 1),               # ICD
-                    dim = c(S, S, 1, n_sim)))
+# rearrange and add homogeneous time dimension
+probs_empty <- array(NA, dim = c(S, S, 1, n_sim))
 
-# rearrange because array() fills by column
-probs <- map(probs, aperm, perm = c(2, 1, 3, 4))
+probs <- list(probs_empty,
+              probs_empty)
+
+probs[[1]][,,1,] <- aperm(lambda.0, c(2,3,1))
+probs[[2]][,,1,] <- aperm(lambda.1, c(2,3,1))
 
 ## run Markov model
[…]

### 2020-11-06 HCM.SCD.CEanalysis/BUGS/script.R
finished rewrite with combined data. need to test

@@ -9,16 +9,25 @@
 
 library(R2jags)
 
-# select:
 data("trans_counts_obs")
 data("trans_counts_risk6")
 
 
-n_S <- 3  # number of states
-J <- 5    # number of time points
+# combine low and high risk
+# into single transition matrix
+empty_mat <- matrix(rep(0,9), nrow = 3)
 
-r.0 <- data_obs$`0`[, c("healthy", "scd_count", "non_scd_count")]
-r.1 <- data_obs$`1`[, c("healthy", "scd_count", "non_scd_count")]
+r.0 <-
+  rbind(
+    cbind(data_obs[[1]][,-4], empty_mat),
+    cbind(empty_mat, data_obs[[2]][,-4]))
+
+r.1 <-
+  rbind(
+    cbind(data_risk6[[1]][, -4], empty_mat),
+    cbind(empty_mat, data_risk6[[2]][, -4]))
+
+n_S <- 6  # number of states
 
 n.0 <- unname(rowSums(r.0))
 n.1 <- unname(rowSums(r.1))
@@ -57,8 +66,6 @@ inits <- function() {
                         rep(NA, n_S)))
 }
 
-inits()
-
 n.iter <- 10000
 n.burnin <- 5000
 n.thin <- floor((n.iter - n.burnin)/500)
@@ -76,14 +83,9 @@ mm1 <-
 
 R2WinBUGS::attach.bugs(mm1$BUGSoutput)
 
-#select:
-save(mm1, file = "data/jags_obs.RData")
-saveRDS(lambda.0, file = "data/lambda0_obs.Rds")
-saveRDS(lambda.1, file = "data/lambda1_obs.Rds")
-
-save(mm1, file = "data/jags_risk6.RData")
-saveRDS(lambda.0, file = "data/lambda0_risk6.Rds")
-saveRDS(lambda.1, file = "data/lambda1_risk6.Rds")
+save(mm1, file = "data/jags_output.RData")
+saveRDS(lambda.0, file = "data/lambda0.Rds")
+saveRDS(lambda.1, file = "data/lambda1.Rds")
 
 # print(mm1, digits = 3, intervals = c(0.025, 0.975))
 

### 2020-11-06 HCM.SCD.CEanalysis/scripts/main_ce_analysis.R

@@ -17,7 +17,10 @@ R2WinBUGS::attach.bugs(mm1$BUGSoutput)
 # 2. SCD
 # 3. all-cause mortality
 
-n_init <- c(1000, 0, 0)
+#TODO: get proportions from dataset
+# pICD <-
+n_init <- c(500, 0, 0, 500, 0, 0)
+
 n_sim <- nrow(lambda.0)
 S <- length(n_init)
 J <- 12  # max time
@@ -27,19 +30,19 @@ J <- 12  # max time
 
 # 0.637: managed with an ICD (Noyes 2007)
 
-e_unit <- list(low_risk = c(1, 0, 0),
-               ICD = c(0.637, 0.5, 0))
+e_unit <- list(obs = c(1, 0, 0, 0.637, 0.5, 0),
+               risk6 = c(1, 0, 0, 0.637, 0.5, 0))
 
 # cost of SCD risk algorithm £20
 # cost of SCD risk management for only day case procedures_new £4792 (Waight 2019)
 # cost of non-fatal HCM related events £22,880 (UK Stroke Association)
 
-c_unit <- list(low_risk = c(0, 0, 0),
-               ICD = c(0, 22880, 0))
+c_unit <- list(obs = c(0, 0, 0, 0, 22880, 0),
+               risk6 = c(0, 0, 0, 0, 22880, 0))
 
 # linear proportion decrease each year in state
-pdecr <- list(low_risk = c(0, 0, 0),
-              ICD = c(0, 0, 0))
+pdecr <- list(obs = c(0, 0, 0, 0, 0, 0),
+              risk6 = c(0, 0, 0, 0, 0, 0))
 
 # rearrange and add homogeneous time dimension
 probs_empty <- array(NA, dim = c(S, S, 1, n_sim))
@@ -50,6 +53,7 @@ probs <- list(probs_empty,
 probs[[1]][,,1,] <- aperm(lambda.0, c(2,3,1))
 probs[[2]][,,1,] <- aperm(lambda.1, c(2,3,1))
 
+
 ## run Markov model
 res_new <-
   init_pop(n_init, n_sim, J) %>%
@@ -58,17 +62,8 @@ res_new <-
          e_unit,
          pdecr)
 
-# total cost for this risk stratification
-total_cost <- res_new$cost[[1]] + res_new$cost[[2]]
-total_eff <- res_new$eff[[1]] + res_new$eff[[2]]
-
[…]

### 2020-11-07 HCM.SCD.CEanalysis/BUGS/script.R
BUGS script inits for new combined trans mat

@@ -15,19 +15,22 @@ data("trans_counts_risk6")
 
 # combine low and high risk
 # into single transition matrix
-empty_mat <- matrix(rep(0,9), nrow = 3)
+# pad with 0s
+empty_mat <- matrix(0,
+                    nrow = nrow(data_obs),
+                    ncol = ncol(data_obs))
 
 r.0 <-
   rbind(
-    cbind(data_obs[[1]][,-4], empty_mat),
-    cbind(empty_mat, data_obs[[2]][,-4]))
+    cbind(data_obs[[1]][, -4], empty_mat),
+    cbind(empty_mat, data_obs[[2]][, -4]))
 
 r.1 <-
   rbind(
     cbind(data_risk6[[1]][, -4], empty_mat),
     cbind(empty_mat, data_risk6[[2]][, -4]))
 
-n_S <- 6  # number of states
+n_S <- nrow(r.0)  # number of states
 
 n.0 <- unname(rowSums(r.0))
 n.1 <- unname(rowSums(r.1))
@@ -50,20 +53,25 @@ params <- c("lambda.0", "lambda.1")
 #
 inits <- function() {
 
-  temp.0 <- rgamma(n_S, scale, 1)
-  sum.temp.0 <- sum(temp.0)
-  p.0 <- temp.0/sum.temp.0
+  matgam.0 <- matrix(rgamma(2*n_S, scale, 1), nrow = 2)
+  sum.matgam.0 <- rowSums(matgam.0)
+  p.0 <- matgam.0/sum.matgam.0
 
-  temp.1 <- rgamma(n_S, scale, 1)
-  sum.temp.1 <- sum(temp.1)
-  p.1 <- temp.1/sum.temp.1
+  matgam.1 <- matrix(rgamma(2*n_S, scale, 1), nrow = 2)
+  sum.matgam.1 <- rowSums(matgam.1)
+  p.1 <- matgam.1/sum.matgam.1
 
-  list(lambda.0 = rbind(p.0,
-                        rep(NA, n_S),
-                        rep(NA, n_S)),
-       lambda.1 = rbind(p.1,
-                        rep(NA, n_S),
-                        rep(NA, n_S)))
+  low_risk <-
+    rbind(p.0,
+          matrix(NA, ncol = n_S, nrow = 4))
+
+  ICD <-
+    rbind(p.1,
+          matrix(NA, ncol = n_S, nrow = 4))
+
[…]

### 2020-11-09 HCM.SCD.CEanalysis/scripts/prep_study_data.R
refactor and run prep_study_data

@@ -36,15 +36,24 @@ fuzzy_noise <-
     rnorm(nrow(data_set1), 0, 0.001)
   } else {0}
 
-# cox risk score > 6% ----
 
-data_set1$risk_over_6 <-
-  as.factor((data_set1$risk_5_years + fuzzy_noise) > 0.06)
+# subset cohort
+data_set1 <-
+  data_set1 %>%
+  mutate(risk_over_6 =
+           data_set1$risk_5_years + fuzzy_noise > 0.06,
+         risk_over_4 =
+           data_set1$risk_5_years + fuzzy_noise > 0.04)
+
+# cox risk score > 6% ----
 
 data_risk6 <-
   data_set1 %>%
-  group_split(risk_over_6) %>%
-  setNames(levels(data_set1$risk_over_6)) %>%
+  mutate(risk_status = ifelse(risk_over_6,
+                              yes = "ICD",
+                              no = "low_risk"),
+         risk_status = as.factor(risk_status)) %>%
+  split(.$risk_status) %>%
   map(.f = annual_trans_counts,
       cycle_length = CYCLE) %>%
   map(.f = obs_aggr_trans_mat)
@@ -55,12 +64,13 @@ save(data_risk6, file = "data/trans_counts_risk6.RData")
 
 # cox risk score > 4% ----
 
-data_set1$risk_over_4 <-
-  as.factor((data_set1$risk_5_years + fuzzy_noise) > 0.04)
-
 data_risk4 <-
-  group_split(data_set1, risk_over_4) %>%
-  setNames(levels(data_set1$risk_over_4)) %>%
+  data_set1 %>%
+  mutate(risk_status = ifelse(risk_over_4,
+                              yes = "ICD",
+                              no = "low_risk"),
+         risk_status = as.factor(risk_status)) %>%
+  split(.$risk_status) %>%
   map(.f = annual_trans_counts,
       cycle_length = CYCLE) %>%
   map(.f = obs_aggr_trans_mat)
@@ -86,8 +96,12 @@ data_set1 <-
          rule_icd = num_rf > rf_threshold)
 
 data_rule <-
-  group_split(data_set1, rule_icd) %>%
-  setNames(unique(data_set1$rule_icd)) %>%
+  data_set1 %>%
+  mutate(risk_status = ifelse(rule_icd,
+                              yes = "ICD",
+                              no = "low_risk"),
+         risk_status = as.factor(risk_status)) %>%
+  split(.$risk_status) %>%
[…]

### 2020-11-09 HCM.SCD.CEanalysis/inst/rmarkdown/draft.Rmd
drafted writeup doc

---
title: "HCM SCD cost-effectiveness analysis"
author: "Nathan Green"
date: "09/11/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

## Data
The main data set is individual-level follow-up data of patients with HCM who may have been given an ICD due to some risk decision.

Health and cost data are obtained from literature and expert opinion.
What values we wish to use will determine the form of the state model.
For instance, if cost are only accrued on entry to a state then we may use a tunnel state.
If cost depend on the patient history then we may need to duplicate a state.

## Methods
The individual-level patient data are first stratified in to two groups for each risk algorithm.
These are

* Partition observed in data set
* Number of risk factors > 1, 2 or 3
* Risk score >6%
* Risk score >4%

We included the option of a fuzzy decision boundary such that near the threshold there is some random variation as to whether a patient received an ICD or not.

This gives us starting state populations for HCM with ICD and HCM without ICD which will be different for each risk decision rule. 
Further, the transition probabilities from these states will differ because of the case mixes.

Using WinBUGS called from R, each new data set is used to generate posterior samples of transition probabilities.

$$x^{(1)}_{i.} \sim \mbox{Multinomial}(p^{(1)}_{i.}, n^{(1)}_i), \;\; i = 1,3$$
$$x^{(2)}_{i.} \sim \mbox{Multinomial}(p^{(2)}_{i.}, n^{(2)}_i), \;\; i = 1,3$$

$$p^{(1)}_{i.} \sim \mbox{Dirichlet}(\boldsymbol{\alpha}^{(1)} ), \;\; i = 1,3$$
$$p^{(2)}_{i.} \sim \mbox{Dirichlet}(\boldsymbol{\alpha}^{(2)} ), \;\; i = 1,3$$

For all other states,

\[
p^{(s)}_{ij} = \left\{
\begin{array}{ll}
1 & \mbox{if $i = j$};\\
0 & \mbox{if $i \neq j$}.
\end{array} \right.
\]

A diagram of the current (undecided) cohort model is given below.

<center>
![](model_diagram.png)
</center>

Depending on what happens to shocked patients the transition matrix may look like the following.

$$
\begin{pmatrix}
p_{11} & p_{12} & 0 & 0 & p_{15}\\
0 & p_{22} & 0 & 0 & p_{25}\\
0 & 0 & p_{33} & p_{34} & p_{35}\\
0 & 0 & 0 & 1 & 0\\
0 & 0 & 0 & 0 & 1
\end{pmatrix}
$$

We have used a year step size but if eg cost are accrued at different intervals then this can be adapted.
The time horizon is set at 12 years but can be changed.

## Workflow

The main files to perform the analysis are:

* `prep_study_data.R` munges the raw data and save the input data in `data/`
* `BUGS/script.R` runs the BUGS code in `BUGS/model.txt`
* `main-ce-analysis.R` performs the cost-effectiveness analysis
* `pop_counts_plot.R` creates output plots

## Results

We give some example output to demonstrate what will be produced with the final model.

Histogram of posterior distributions for state transition probabilities.

![](posterior_histograms.png)

Below is an example of state occupancy over time plot.
This shows that for the new algorithm there are fewer scd and more shocks.

![](../../images/state_pop_over_time.png)

### 2020-11-09 HCM.SCD.CEanalysis/BUGS/script.R

@@ -17,18 +17,18 @@ data("trans_counts_risk6")
 # into single transition matrix
 # pad with 0s
 empty_mat <- matrix(0,
-                    nrow = nrow(data_obs),
-                    ncol = ncol(data_obs))
+                    nrow = nrow(data_obs$ICD),
+                    ncol = nrow(data_obs$ICD))
 
 r.0 <-
   rbind(
-    cbind(data_obs[[1]][, -4], empty_mat),
-    cbind(empty_mat, data_obs[[2]][, -4]))
+    cbind(data_obs$ICD[, -4], empty_mat),
+    cbind(empty_mat, data_obs$low_risk[, -4]))
 
 r.1 <-
   rbind(
-    cbind(data_risk6[[1]][, -4], empty_mat),
-    cbind(empty_mat, data_risk6[[2]][, -4]))
+    cbind(data_risk6$ICD[, -4], empty_mat),
+    cbind(empty_mat, data_risk6$low_risk[, -4]))
 
 n_S <- nrow(r.0)  # number of states
 
@@ -48,7 +48,7 @@ dataJags <-
        n_S = n_S)
 
 filein <- "BUGS/model.txt"
-params <- c("lambda.0", "lambda.1")
+params <- c("lambda.0", "lambda.1") #probabilities
 
 #
 inits <- function() {

### 2020-11-09 HCM.SCD.CEanalysis/R/obs_aggr_trans_mat.R

@@ -10,7 +10,7 @@
 #' @export
 #'
 obs_aggr_trans_mat <- function(data,
-                               max_year = 5) {
+                               max_year = 20) {
 
   trans_mat <- list()
   state_names <- c("healthy", "scd_count", "non_scd_count")

### 2020-11-09 HCM.SCD.CEanalysis/scripts/decision_tree.R

@@ -1,26 +1,0 @@
-
-#
-#
-#
-
-c_ICD <- 1
-p_ICD_under4
-p_ICD_4to6
-p_ICD_over6
-p_under4
-p_4to6
-p_over6
-
-# c_assess + c_ICD * (p_under4*p_ICD_under4 + p_4to6*p_ICD_4to6 + p_over6*p_ICD_over6)
-#
-# and values we need are:
-#
-# cost of assessment (c_assess)
-# cost of ICD (c_ICD)
-# probability assessed <4%, 4-6%, >6%
-# probability ICD given assessment result
-# Alive with HCM: unit cost, unit utility for ICD and non-ICD
-# p(alive with HCM -> SCD) for ICD and non-ICD, by age (or other covariate?)
-# p(alive with HCM -> all cause death) for ICD and non-ICD, by age (or other covariate?)
-
-

### 2020-11-09 HCM.SCD.CEanalysis/scripts/main_ce_analysis.R

@@ -7,10 +7,10 @@
 library(dplyr)
 library(purrr)
 library(BCEA)
-library(res_newhape2)
+library(reshape2)
 library(HCM.SCD.CEanalysis)
 
-data(mm1)
+data("jags_output")
 R2WinBUGS::attach.bugs(mm1$BUGSoutput)
 
 # 1. alive with HCM
@@ -25,23 +25,24 @@ n_sim <- nrow(lambda.0)
 S <- length(n_init)
 J <- 12  # max time
 
-# utilities and costs
+# utilities and costs ----
 # for each intervention
 
 # 0.637: managed with an ICD (Noyes 2007)
 
-e_unit <- list(obs = c(1, 0, 0, 0.637, 0.5, 0),
-               risk6 = c(1, 0, 0, 0.637, 0.5, 0))
+e_unit <- list(obs =   c(0.637, 0, 0, 1, 0.5, 0),
+               risk6 = c(0.637, 0, 0, 1, 0.5, 0))
 
-# cost of SCD risk algorithm £20
-# cost of SCD risk management for only day case procedures_new £4792 (Waight 2019)
 # cost of non-fatal HCM related events £22,880 (UK Stroke Association)
+# cost of SCD risk management for only day case procedures_new £4792 (Waight 2019)
 
-c_unit <- list(obs = c(0, 0, 0, 0, 22880, 0),
-               risk6 = c(0, 0, 0, 0, 22880, 0))
+##TODO: do we need a tunnel state for shock so that its a one-off cost?
+
+c_unit <- list(obs =   c(0, 0, 22880, 0, 0, 0),
+               risk6 = c(0, 0, 22880, 0, 0, 0))
 
 # linear proportion decrease each year in state
-pdecr <- list(obs = c(0, 0, 0, 0, 0, 0),
+pdecr <- list(obs =   c(0, 0, 0, 0, 0, 0),
               risk6 = c(0, 0, 0, 0, 0, 0))
 
 # rearrange and add homogeneous time dimension
@@ -65,7 +66,11 @@ res_new <-
 c <- map_dfc(res_new$cost, rowSums) %>% as.matrix()
 e <- map_dfc(res_new$eff, rowSums) %>% as.matrix()
 
-labels <- c("low_risk", "ICD")
+
+##TODO: add one-off cost
+# cost of SCD risk algorithm £20
+
+labels <- c("obs", "risk6")
 
 m <-
   BCEA::bcea(e, c,

### 2020-11-09 HCM.SCD.CEanalysis/scripts/pop_counts_plot.R

@@ -1,0 +1,71 @@
+
+# pop counts plot
+# stacked percentage bar plot
+
+
+library(ggplot2)
+library(gridExtra)
+
+plot_obs <-
+  apply(res_new$pop[[1]], c(1,2), mean) %>%
+  as.data.frame()
+
+plot_obs$state <-
+  c("ICD", "shock", "all_cause", "low_risk", "scd", "all_cause")
+
+plot_obs <-
+  melt(plot_obs,
+       id.vars = "state",
+       value.name = "pop",
+       variable.name = "time") %>%
+  group_by(state, time) %>%
+  summarise(pop = sum(pop))
+
+plot_obs$time <- as.numeric(gsub("V", "", plot_obs$time))
+plot_obs$state <-
+  factor(plot_obs$state,
+         levels = c("scd", "shock", "all_cause", "ICD", "low_risk"))
+
+p1 <-
+  ggplot(plot_obs, aes(fill = state, y = pop, x = time)) +
+  geom_bar(position = "fill", stat = "identity") +
+  scale_x_continuous(breaks = 1:12)
+
+plot_risk6 <-
+  apply(res_new$pop[[2]], c(1,2), mean) %>%
+  as.data.frame()
+
+plot_risk6$state <-
+  c("ICD", "shock", "all_cause", "low_risk", "scd", "all_cause")
+
+plot_risk6 <-
+  melt(plot_risk6,
+       id.vars = "state",
+       value.name = "pop",
+       variable.name = "time") %>%
+  group_by(state, time) %>%
+  summarise(pop = sum(pop))
+
+plot_risk6$time <- as.numeric(gsub("V", "", plot_risk6$time))
+plot_risk6$state <-
+  factor(plot_risk6$state,
+         levels = c("scd", "shock", "all_cause", "ICD", "low_risk"))
+
+p2 <-
+  ggplot(plot_risk6, aes(fill = state, y = pop, x = time)) +
+  geom_bar(position = "fill", stat = "identity") +
+  scale_x_continuous(breaks = 1:12)
+
+# combine
+total_dat <- rbind(cbind(strat = "obs", plot_obs),
+                   cbind(strat = "risk6", plot_risk6))
+
+facet_plot <-
+  ggplot(total_dat, aes(fill = state, y = pop, x = time)) +
+  geom_bar(position = "fill", stat = "identity") +
+  scale_x_continuous(breaks = 1:12) +
+  facet_grid(.~strat)
+
[…]

### 2020-11-09 HCM.SCD.CEanalysis/scripts/posterior_hist_plots.R

@@ -1,0 +1,30 @@
+
+# posterior histograms
+# transition probabilities
+#
+
+
+data("jags_output")
+R2WinBUGS::attach.bugs(mm1$BUGSoutput)
+
+x11()
+par(mfrow = c(2,3))
+hist(lambda.0[, 1, 1], breaks = 20, xlab = "Prob", main = "ICD -> ICD")
+abline(v = mean(lambda.0[, 1, 1]), col = "red")
+hist(lambda.0[, 1, 2], breaks = 20, xlab = "Prob", main = "ICD -> Shock")
+abline(v = mean(lambda.0[, 1, 2]), col = "red")
+hist(lambda.0[, 1, 3], breaks = 20, xlab = "Prob", main = "ICD -> All-cause death")
+abline(v = mean(lambda.0[, 1, 3]), col = "red")
+
+hist(lambda.0[, 4, 4], breaks = 20, xlab = "Prob", main = "Low risk -> Low risk")
+abline(v = mean(lambda.0[, 4, 4]), col = "red")
+hist(lambda.0[, 4, 5], breaks = 20, xlab = "Prob", main = "Low risk -> SCD")
+abline(v = mean(lambda.0[, 4, 5]), col = "red")
+hist(lambda.0[, 4, 6], breaks = 20, xlab = "Prob", main = "Low risk -> All-cause death")
+abline(v = mean(lambda.0[, 4, 6]), col = "red")
+
+
+
+
+
+

### 2020-11-09 HCM.SCD.CEanalysis/scripts/prep_study_data.R

@@ -13,6 +13,9 @@
 # then aggregated across some time horizon and formatted
 # as a transition matrix
 
+##TODO: what period to take obs_aggr_trans_mat() over?
+##      at the moment 5 years only
+
 
 library(haven)
 library(dplyr)

### 2020-11-12 HCM.SCD.CEanalysis/BUGS/script.R
added certain transition from shock to HCM ICD

@@ -5,6 +5,14 @@
 # estimate transition probabilities
 # from discrete time counts
 # using multinomial likelihood
+#
+# transition matrix:
+# 1 ICD      x x x 0 0 0
+# 2 shock    1 0 0 0 0 0
+# 3 death    0 0 1 0 0 0
+# 4 low risk 0 0 0 x x x
+# 5 scd      0 0 0 0 1 0
+# 6 death    0 0 0 0 0 1
 
 
 library(R2jags)
@@ -13,39 +21,38 @@ data("trans_counts_obs")
 data("trans_counts_risk6")
 
 
+n_S <- nrow(data_obs$ICD)  # number of states
+
 # combine low and high risk
 # into single transition matrix
 # pad with 0s
-empty_mat <- matrix(0,
-                    nrow = nrow(data_obs$ICD),
-                    ncol = nrow(data_obs$ICD))
+empty_mat <- matrix(0, nrow = n_S, ncol = n_S)
 
 r.0 <-
   rbind(
-    cbind(data_obs$ICD[, -4], empty_mat),
-    cbind(empty_mat, data_obs$low_risk[, -4]))
+    cbind(data_obs$ICD[, 1:n_S], empty_mat),
+    cbind(empty_mat, data_obs$low_risk[, 1:n_S]))
 
 r.1 <-
   rbind(
-    cbind(data_risk6$ICD[, -4], empty_mat),
-    cbind(empty_mat, data_risk6$low_risk[, -4]))
-
-n_S <- nrow(r.0)  # number of states
+    cbind(data_risk6$ICD[, 1:n_S], empty_mat),
+    cbind(empty_mat, data_risk6$low_risk[, 1:n_S]))
 
 n.0 <- unname(rowSums(r.0))
 n.1 <- unname(rowSums(r.1))
 
 scale <- 1                            # level of informativeness for
-alpha.0 <- alpha.1 <- rep(scale, n_S) # the Dirichlet prior
+alpha.0 <- alpha.1 <- rep(scale, 2*n_S) # the Dirichlet prior
 
 dataJags <-
   list(n.0 = n.0,
        n.1 = n.1,
        r.0 = r.0,
        r.1 = r.1,
+       from_shock = c(1,0,0,0,0,0),
+       from_ICD_death = c(0,0,1,0,0,0),
        alpha.0 = alpha.0,
[…]

### 2020-11-12 HCM.SCD.CEanalysis/R/obs_aggr_trans_mat.R

@@ -4,6 +4,8 @@
 #' Create aggregate observed transition counts
 #' see p.185 Table 5.2 chapter 5 in BCEA book
 #'
+#' Assume 3 state model: "healthy", "scd_count", "non_scd_count"
+#'
 #' @param data new state arrivals at time step
 #'             e.g. data = data_rule$`FALSE`
 #' @param max_year

### 2020-11-12 HCM.SCD.CEanalysis/scripts/prep_study_data.R

@@ -13,7 +13,7 @@
 # then aggregated across some time horizon and formatted
 # as a transition matrix
 
-##TODO: what period to take obs_aggr_trans_mat() over?
+##TODO: what periodl to take obs_aggr_trans_mat() over?
 ##      at the moment 5 years only
 
 

### 2020-11-12 HCM.SCD.CEanalysis/inst/rmarkdown/draft.Rmd
use actual starting populations in data for ce analysis and modify init_pop() to take a list

@@
 ---
-title: "HCM SCD cost-effectiveness analysis"
+title: "HCM SCD risk algorithm cost-effectiveness analysis"
 author: "Nathan Green"
@@
 output: pdf_document
+editor_options: 
+  markdown: 
+    wrap: sentence
+references:
+- id: mahony2014
+  title: A novel clinical risk prediction model for sudden cardiac death in hypertrophic cardiomyopathy (HCM risk-SCD)
+  author:
+  - family: O'Mahony
+    given: C
+  container-title: Eur Heart
+  volume: 35
+  URL: 'https://pubmed.ncbi.nlm.nih.gov/24126876/'
+  DOI:
+  issue: 30
+  page: 2010-20
+  type: article-journal
+  issued:
+    year: 2014
 ---
@@
 
+Risk score algorithm details see [@mahony2014].
+
 ## Data
-The main data set is individual-level follow-up data of patients with HCM who may have been given an ICD due to some risk decision.
 
-Health and cost data are obtained from literature and expert opinion.
+The main data set contains individual-level follow-up data of patients with HCM who may have been given an ICD due to some risk decision.
+
+Health and cost data were obtained from literature and expert opinion.
 What values we wish to use will determine the form of the state model.
@@
 
+Table 1 gives the unit cost and health values used in the model.
+
+ Name | Value | Parameter | Source 
+------|-------|-----------|--------
+| _Health_ ||||
+| Manage with ICD | 0.637 QALY/year | $h_1$ | Noyes 2007 |
+| Shock | ? utility losss | $h_2$ |  |
+| HCM | ? QALY/year | $h_3$ |  |
+| Death | 0 QALY/year | $h_4$, $h_5$ |  |
+|       |      |      |     |
+| _Cost_ ||||
+| Manage with ICD | ? | $c_1$ |  |
+| Non-fatal HCM event | £22,880 | $c_2$ | UK Stroke Assoc. |
+| HCM | ? | $c_3$ |  |
+| SCD | ? | $c_4$ |  |
+| All-cause death | ? | $c_5$ |  |
+|      |      |      |     |
+
+: Model parameter values.
+
 ## Methods
+
 The individual-level patient data are first stratified in to two groups for each risk algorithm.
@@
 
-* Partition observed in data set
-* Number of risk factors > 1, 2 or 3
-* Risk score >6%
-* Risk score >4%
+-   Partition observed in data set
+-   ICD given if number of risk factors \> 1, 2 or 3
+-   ICD given if risk score \>6%
+-   ICD given if risk score \>4%
 
@@
 
-Using WinBUGS called from R, each new data set is used to generate posterior samples of transition probabilities.
+A diagram of the current cohort model is given below.
 
-$$x^{(1)}_{i.} \sim \mbox{Multinomial}(p^{(1)}_{i.}, n^{(1)}_i), \;\; i = 1,3$$
-$$x^{(2)}_{i.} \sim \mbox{Multinomial}(p^{(2)}_{i.}, n^{(2)}_i), \;\; i = 1,3$$
+```{r fig.align="center", echo=FALSE}
+knitr::include_graphics("model_diagram.png")
+```
 
-$$p^{(1)}_{i.} \sim \mbox{Dirichlet}(\boldsymbol{\alpha}^{(1)} ), \;\; i = 1,3$$
-$$p^{(2)}_{i.} \sim \mbox{Dirichlet}(\boldsymbol{\alpha}^{(2)} ), \;\; i = 1,3$$
-
-For all other states,
-
-\[
-p^{(s)}_{ij} = \left\{
-\begin{array}{ll}
-1 & \mbox{if $i = j$};\\
-0 & \mbox{if $i \neq j$}.
-\end{array} \right.
-\]
-
-A diagram of the current (undecided) cohort model is given below.
-
-<center>
-![](model_diagram.png)
-</center>
-
-Depending on what happens to shocked patients the transition matrix may look like the following.
+Assuming that shocked patients return to the HCM ICD state then the transition matrix may look like the following.
 
@@
 p_{11} & p_{12} & 0 & 0 & p_{15}\\
-0 & p_{22} & 0 & 0 & p_{25}\\
+1 & 0 & 0 & 0 & 0\\
 0 & 0 & p_{33} & p_{34} & p_{35}\\
@@
 
+#### Transition probability inference
+
+Using WinBUGS called from R, each new data set is used to generate posterior samples of transition probabilities.
+Denote $x$ as the observed number of transitions, $p$ the probability of a transition and $n$ as the total number of transitions from a given state.
+The hyperparameters $\alpha$ characterise the prior knowledge on $p$.
+Superscripts indicate the decision rule used.
+
+$$x^{(1)}_{i.} \sim \mbox{Multinomial}(p^{(1)}_{i.}, n^{(1)}_i), \;\; i = 1,3$$
+$$x^{(2)}_{i.} \sim \mbox{Multinomial}(p^{(2)}_{i.}, n^{(2)}_i), \;\; i = 1,3$$
+
+$$p^{(1)}_{i.} \sim \mbox{Dirichlet}(\boldsymbol{\alpha}^{(1)} ), \;\; i = 1,3$$
+$$p^{(2)}_{i.} \sim \mbox{Dirichlet}(\boldsymbol{\alpha}^{(2)} ), \;\; i = 1,3$$
+
+For all sink states,
+
+$$
+p^{(s)}_{ij} = \left\{
+\begin{array}{ll}
+1 & \mbox{if $i = j$};\\
+0 & \mbox{if $i \neq j$}.
+\end{array} \right.
+$$
+
 ## Workflow
@@
 
-* `prep_study_data.R` munges the raw data and save the input data in `data/`
-* `BUGS/script.R` runs the BUGS code in `BUGS/model.txt`
-* `main-ce-analysis.R` performs the cost-effectiveness analysis
-* `pop_counts_plot.R` creates output plots
+-   `prep_study_data.R` munges the raw data and save the input data in `data/`
+-   `BUGS/script.R` runs the BUGS code in `BUGS/model.txt`
+-   `main-ce-analysis.R` performs the cost-effectiveness analysis
+-   `pop_counts_plot.R` creates output plots
 
@@
 ![](../../images/state_pop_over_time.png)
+
+# References

### 2020-11-12 HCM.SCD.CEanalysis/R/init_pop.R

@@ -1,10 +1,11 @@
 
 #' Create initial state population
 #'
-#' @param pop Number of individuals (state, time, n_sim)
+#' @param n_init Initial state populations; scalar or list
 #' @param n_sim Number of simulations (from jags)
 #' @param tmax Time horizon
 #' @param n_interv Number of interventions, including status-quo
+#'
 #' @return List [[interv]] state x time x sim
 #' @seealso [ce_sim()]
 #'
@@ -15,16 +16,24 @@ init_pop <- function(n_init,
                      tmax,
                      n_interv = 2) {
 
-  S <- length(n_init)
+  if (is.list(n_init) & length(n_init) != n_interv)
+    stop("n_interv and length of n_init don't match", call. = FALSE)
+
+  if (length(n_init) == 1)
+    n_init <- purrr::map(1:n_interv, ~n_init)
+
+  S <- length(n_init[[1]])
   pop <- list()
 
   for (n in seq_len(n_interv)) {
     pop[[n]] <- array(NA, c(S, tmax, n_sim))
 
     for (i in seq_len(n_sim)) {
-      pop[[n]][, 1, i] <- n_init
+      pop[[n]][, 1, i] <- n_init[[n]]
     }
   }
 
   pop
 }
+
+

### 2020-11-12 HCM.SCD.CEanalysis/scripts/main_ce_analysis.R

@@ -13,13 +13,25 @@ library(HCM.SCD.CEanalysis)
 data("jags_output")
 R2WinBUGS::attach.bugs(mm1$BUGSoutput)
 
-# 1. alive with HCM
-# 2. SCD
-# 3. all-cause mortality
+# individual patient data
+data("ipd_risk")
 
-#TODO: get proportions from dataset
-# pICD <-
-n_init <- c(500, 0, 0, 500, 0, 0)
+# 1. HCM with ICD
+# 2. shock
+# 3. all-cause mortality
+# 4. HCM
+# 5. SCD
+# 6. all-cause mortality
+
+# start state populations
+init_risk6 <- table(ipd_risk$risk_over_6)
+init_obs <- table(ipd_risk$icd)
+
+n_init <- list()
+n_init$obs <- c(init_obs["1"], 0, 0,         # ICD
+                init_obs["0"], 0, 0)         # low risk
+n_init$risk6 <- c(init_risk6["TRUE"], 0, 0,
+                  init_risk6["FALSE"], 0, 0)
 
 n_sim <- nrow(lambda.0)
 S <- length(n_init)

### 2020-11-12 HCM.SCD.CEanalysis/scripts/prep_study_data.R

@@ -13,9 +13,6 @@
 # then aggregated across some time horizon and formatted
 # as a transition matrix
 
-##TODO: what periodl to take obs_aggr_trans_mat() over?
-##      at the moment 5 years only
-
 
 library(haven)
 library(dplyr)
@@ -41,17 +38,20 @@ fuzzy_noise <-
 
 
 # subset cohort
-data_set1 <-
+ipd_risk <-
   data_set1 %>%
   mutate(risk_over_6 =
            data_set1$risk_5_years + fuzzy_noise > 0.06,
          risk_over_4 =
            data_set1$risk_5_years + fuzzy_noise > 0.04)
 
+save(ipd_risk, file = "data/ipd_risk.RData")
+
+
 # cox risk score > 6% ----
 
 data_risk6 <-
-  data_set1 %>%
+  ipd_risk %>%
   mutate(risk_status = ifelse(risk_over_6,
                               yes = "ICD",
                               no = "low_risk"),
@@ -68,7 +68,7 @@ save(data_risk6, file = "data/trans_counts_risk6.RData")
 # cox risk score > 4% ----
 
 data_risk4 <-
-  data_set1 %>%
+  ipd_risk %>%
   mutate(risk_status = ifelse(risk_over_4,
                               yes = "ICD",
                               no = "low_risk"),
@@ -94,7 +94,7 @@ save(data_risk4, file = "data/trans_counts_risk4.RData")
 rf_threshold <- 1
 
 data_set1 <-
-  data_set1 %>%
+  ipd_risk %>%
   mutate(num_rf = mwt30 + nsvt + fhxscd + syncope,
          rule_icd = num_rf > rf_threshold)
 
@@ -118,7 +118,7 @@ save(data_rule, file = "data/trans_counts_rule.RData")
 # i.e. observed ICD implants
 
 data_obs <-
-  data_set1 %>%
+  ipd_risk %>%
   mutate(risk_status = ifelse(icd,
                               yes = "ICD",
                               no = "low_risk"),

### 2020-11-27 HCM.SCD.CEanalysis/inst/rmarkdown/draft.Rmd
fix structural seros in markov model Dirichlet. add starting state costs.

@@
 
-Risk score algorithm details see [@mahony2014].
+Sudden cardiac death (SCD) risk score algorithm details see [@mahony2014].
 
@@
 
- Name | Value | Parameter | Source 
+ Name | Value$^*$ | Parameter | Source 
 ------|-------|-----------|--------
@@
 | Manage with ICD | 0.637 QALY/year | $h_1$ | Noyes 2007 |
-| Shock | ? utility losss | $h_2$ |  |
-| HCM | ? QALY/year | $h_3$ |  |
+| Shock | one-off and/or on-going? utility loss | $h_2$ |  |
+| HCM without ICD | < 1 ? QALY/year | $h_3$ |  |
 | Death | 0 QALY/year | $h_4$, $h_5$ |  |
@@
 | Manage with ICD | ? | $c_1$ |  |
+| Perform risk score | £20? |  |  |
+| Implant ICD | £4,666? |  | EY02B Tariffs |
+| Attention to ICD | £1,817? |  | EY10B Tariffs |
+| Testing ICD | £467? |  | Tariffs |
 | Non-fatal HCM event | £22,880 | $c_2$ | UK Stroke Assoc. |
-| HCM | ? | $c_3$ |  |
+| e.g shock? | one-off and/or on-going? | | |
+| HCM without ICD | ? | $c_3$ |  |
 | SCD | ? | $c_4$ |  |
@@
 
-: Model parameter values.
+: Model parameter values. $^*$either one-off/on state entry or recurring.
+
+Table 2 gives the starting state populations for non-zero states.
+
+Risk rule | State | Population
+----------|-------|-----------
+Observed  | HCM ICD | 559
+      -    | HCM | 3113
+Score > 4%| HCM ICD | 2561
+      -    | HCM | 1111
+Score > 6%| HCM ICD | 542
+      -    | HCM | 3130
+Risk factor >0| HCM ICD | 1785
+      -    | HCM | 1887
+Risk factor >1| HCM ICD | 481
+      -    | HCM | 3191
+Risk factor >2| HCM ICD | 78
+      -    | HCM | 3594
 
@@
 -   Partition observed in data set
--   ICD given if number of risk factors \> 1, 2 or 3
+-   ICD given if number of risk factors \> 0, 1 or 2
 -   ICD given if risk score \>6%

### 2020-11-27 HCM.SCD.CEanalysis/BUGS/script.R

@@ -41,8 +41,9 @@ r.1 <-
 n.0 <- unname(rowSums(r.0))
 n.1 <- unname(rowSums(r.1))
 
-scale <- 1                            # level of informativeness for
-alpha.0 <- alpha.1 <- rep(scale, 2*n_S) # the Dirichlet prior
+scale <- 1                                 # level of informativeness for
+alpha.0 <- c(rep(scale, n_S), rep(0, n_S)) # the Dirichlet prior
+alpha.1 <- c(rep(0, n_S), rep(scale, n_S)) # with structural zeros
 
 dataJags <-
   list(n.0 = n.0,
@@ -60,21 +61,36 @@ params <- c("lambda.0", "lambda.1") #probabilities
 #
 inits <- function() {
 
-  matgam.0 <- matrix(rgamma(2*2*n_S, scale, 1), nrow = 2)
-  sum.matgam.0 <- rowSums(matgam.0)
-  p.0 <- matgam.0/sum.matgam.0
+  matgam.0 <- matrix(rgamma(2*2*n_S, scale, 1),
+                     nrow = 2)
+  # prohibited transitions
+  matgam.0[1, (n_S+1):(2*n_S)] <- 0
+  matgam.0[2, 1:n_S] <- 0
 
-  matgam.1 <- matrix(rgamma(2*2*n_S, scale, 1), nrow = 2)
-  sum.matgam.1 <- rowSums(matgam.1)
+  sum.matgam.0 <- rowSums(matgam.0, na.rm = TRUE)
+
+  matgam.1 <- matrix(rgamma(2*2*n_S, scale, 1),
+                     nrow = 2)
+  # prohibited transitions
+  matgam.1[1, (n_S+1):(2*n_S)] <- 0
+  matgam.1[2, 1:n_S] <- 0
+
+  sum.matgam.1 <- rowSums(matgam.1, na.rm = TRUE)
+
+  p.0 <- matgam.0/sum.matgam.0
   p.1 <- matgam.1/sum.matgam.1
 
   obs_mat <-
     rbind(p.0,
-          matrix(NA, ncol = 2*n_S, nrow = 2*n_S - 2))
+          matrix(NA,
+                 ncol = 2*n_S,
+                 nrow = 2*n_S - 2))
 
   risk6_mat <-
     rbind(p.1,
-          matrix(NA, ncol = 2*n_S, nrow = 2*n_S - 2))
+          matrix(NA,
[…]

### 2020-11-27 HCM.SCD.CEanalysis/R/annual_trans_counts.R

@@ -29,6 +29,8 @@ annual_trans_counts <- function(data,
            cum_non_scd = cumsum(non_scd_count),
            cum_cens = cumsum(cens_count),
            healthy = n_pop - cum_cens - cum_scd - cum_non_scd,
-           at_risk = lag(healthy, default = n_pop))
+           at_risk0 = lag(healthy, default = n_pop),
+           # actuarial assumption
+           at_risk = round(at_risk0 - 0.5*cens_count, 0))
 }
 

### 2020-11-27 HCM.SCD.CEanalysis/R/ce_sim.R

@@ -1,10 +1,11 @@
 
 #' Markov model cost-effectiveness simulation
 #'
-#' @param pop Number of individuals. List-matrix [[n_int]] n_sim x time
+#' @param pop Number of individuals. See `init_pop()`. List-matrix [[n_int]] n_sim x time
 #' @param probs Transition probabilities. List-array [[n_int]] states x states x time x sim
 #' @param c_unit Unit costs per state. List [[n_int]]
 #' @param e_unit Health unit values per state. List [[n_int]]
+#' @param c_init One-off initial state costs. List [[n_int]]
 #' @param pdecr Linear utility decrease per state
 #' @param delta Discount rate; default 3.5\% as 0.035
 #' @return List
@@ -18,6 +19,7 @@ ce_sim <- function(pop,
                    probs,
                    c_unit,
                    e_unit,
+                   c_init = NA,
                    pdecr = NA,
                    delta = 0.035) {
 
@@ -66,6 +68,12 @@ ce_sim <- function(pop,
         dcost[[k]][i, j] <- cost[[k]][i, j] / disc
         deff[[k]][i, j] <- eff[[k]][i, j] / disc
       }
+    }
+
+    if (!any(is.na(c_init))) {
+      # add one-off starting state costs
+      cost[[k]][, 1] <- dcost[[k]][, 1] <-
+        c_init[[k]] %*% pop[[k]][,1,1]
     }
   }
 

### 2020-11-27 HCM.SCD.CEanalysis/scripts/main_ce_analysis.R

@@ -1,7 +1,7 @@
 
-# Markov model
+# Markov model script
 # main_ce_analysis.R
-# N Green
+# N Green, UCL
 
 
 library(dplyr)
@@ -34,24 +34,56 @@ n_init$risk6 <- c(init_risk6["TRUE"], 0, 0,
                   init_risk6["FALSE"], 0, 0)
 
 n_sim <- nrow(lambda.0)
-S <- length(n_init)
+S <- length(n_init$obs)
 J <- 12  # max time
 
 # utilities and costs ----
 # for each intervention
 
-# 0.637: managed with an ICD (Noyes 2007)
+# managed with an ICD (Noyes 2007)
+u_hcm <- 0.637
 
-e_unit <- list(obs =   c(0.637, 0, 0, 1, 0.5, 0),
-               risk6 = c(0.637, 0, 0, 1, 0.5, 0))
+##TODO: what is the detriment to having an icd and scd?
+u_icd <- -0.05
+u_scd <- 0.5
+
+e_unit <- list(obs =   c(u_hcm + u_icd, u_scd, 0,
+                         u_hcm,         0,     0),
+               risk6 = c(u_hcm + u_icd, u_scd, 0,
+                         u_hcm,         0,     0))
 
 # cost of non-fatal HCM related events £22,880 (UK Stroke Association)
-# cost of SCD risk management for only day case procedures_new £4792 (Waight 2019)
+c_nfatal <- 22880
 
-##TODO: do we need a tunnel state for shock so that its a one-off cost?
+##TODO: fatal cost?
 
-c_unit <- list(obs =   c(0, 0, 22880, 0, 0, 0),
-               risk6 = c(0, 0, 22880, 0, 0, 0))
+# cost of SCD risk management for day case procedures_new £4792 (Waight 2019)
+
+# EY02B NHS tariffs implantation cost
+c_icd <- 4666
+c_rscore <- 20
+
+c_init <- list(obs =   c(c_icd, 0, 0,
+                         0, 0, 0),
+               risk6 = c(c_icd + c_rscore, 0, 0,
+                         0, 0, 0))
+
+##TODO: for shock so that its a one-off cost: extend code or tunnel state?
+
+# what complexity & comorbidity (CC) score?
+# other possible cost from NHS tariffs:
+# * attention to icd £1817
+# * testing icd £467
+
+c_unit <- list(obs =   c(0, 0, c_nfatal,
+                         0, 0, 0),
[…]

### 2020-11-27 HCM.SCD.CEanalysis/scripts/prep_study_data.R

@@ -38,12 +38,22 @@ fuzzy_noise <-
 
 
 # subset cohort
+rf_threshold <- 1
+
 ipd_risk <-
   data_set1 %>%
   mutate(risk_over_6 =
            data_set1$risk_5_years + fuzzy_noise > 0.06,
          risk_over_4 =
-           data_set1$risk_5_years + fuzzy_noise > 0.04)
+           data_set1$risk_5_years + fuzzy_noise > 0.04,
+         num_rf = mwt30 + nsvt + fhxscd + syncope,
+         rule_icd = num_rf > rf_threshold)
+
+# mwt30: max wall thickness
+# nsvt: NSVT
+# fhxscd: family history of SCD
+# syncope: unexplained syncope
+# no noise on decision
 
 save(ipd_risk, file = "data/ipd_risk.RData")
 
@@ -83,23 +93,9 @@ data_risk4
 save(data_risk4, file = "data/trans_counts_risk4.RData")
 
 # status-quo as risk factor rule ----
-#
-# mwt30: max wall thickness
-# nsvt: NSVT
-# fhxscd: family history of SCD
-# syncope: unexplained syncope
-#
-# no noise on decision
-
-rf_threshold <- 1
-
-data_set1 <-
-  ipd_risk %>%
-  mutate(num_rf = mwt30 + nsvt + fhxscd + syncope,
-         rule_icd = num_rf > rf_threshold)
 
 data_rule <-
-  data_set1 %>%
+  ipd_risk %>%
   mutate(risk_status = ifelse(rule_icd,
                               yes = "ICD",
                               no = "low_risk"),

### 2021-04-20 HCM.SCD.CEanalysis/inst/rmarkdown/draft.Rmd
replace constants cost inputs with functions to sample costs at runtime

--- 2020-11-12 HCM.SCD.CEanalysis/inst/rmarkdown/draft.Rmd
@@
     wrap: sentence
-references:
-- id: mahony2014
-  title: A novel clinical risk prediction model for sudden cardiac death in hypertrophic cardiomyopathy (HCM risk-SCD)
-  author:
-  - family: O'Mahony
-    given: C
-  container-title: Eur Heart
-  volume: 35
-  URL: 'https://pubmed.ncbi.nlm.nih.gov/24126876/'
-  DOI:
-  issue: 30
-  page: 2010-20
-  type: article-journal
-  issued:
-    year: 2014
+bibliography: bibliography.bib
 ---
@@
 
-Risk score algorithm details see [@mahony2014].
+Sudden cardiac death (SCD) risk score algorithm details see [@OMahony2014].
+
+There have been previous cost-effectiveness analyses [@Magnusson2020; @Boriani2014; @Bryant2007; @Caro2007; @Cowie2009; @Garcia-Perez2015; @Mealing2016; @Sanders2005; @Smith2013]
 
@@
 
- Name | Value | Parameter | Source 
-------|-------|-----------|--------
-| _Health_ ||||
-| Manage with ICD | 0.637 QALY/year | $h_1$ | Noyes 2007 |
-| Shock | ? utility losss | $h_2$ |  |
-| HCM | ? QALY/year | $h_3$ |  |
-| Death | 0 QALY/year | $h_4$, $h_5$ |  |
++----------------------------------+-----------------+--------+------------------+
+| Description                      | Value$^*$       | Range  | Source           |
++==================================+=================+========+==================+
+| *Health*                         |                 |        |                  |
++----------------------------------+-----------------+--------+------------------+
+| Manage with ICD                  | 0.637 QALY/year |        | @Noyes2007       |
++----------------------------------+-----------------+--------+------------------+
+| Implantation procedure utility   | -0.016          |        | @Smith2013       |
++----------------------------------+-----------------+--------+------------------+
+| Shock utility                    | -0.5            |        |                  |
++----------------------------------+-----------------+--------+------------------+
+| HCM without ICD                  | 1 QALY/year     |        |                  |
++----------------------------------+-----------------+--------+------------------+
+| Death                            | 0 QALY/year     |        |                  |
++----------------------------------+-----------------+--------+------------------+
 |                                  |                 |        |                  |
-| _Cost_ ||||
-| Manage with ICD | ? | $c_1$ |  |
-| Non-fatal HCM event | £22,880 | $c_2$ | UK Stroke Assoc. |
-| HCM | ? | $c_3$ |  |
-| SCD | ? | $c_4$ |  |
-| All-cause death | ? | $c_5$ |  |
++----------------------------------+-----------------+--------+------------------+
+| *Cost*                           |                 |        |                  |
++----------------------------------+-----------------+--------+------------------+
+| ICD appointment                  | £10             |        |                  |
++----------------------------------+-----------------+--------+------------------+
+| Perform risk score               | £20?            |        |                  |
++----------------------------------+-----------------+--------+------------------+
+| Implant ICD                      | £4,666          |        | EY02B Tariffs    |
++----------------------------------+-----------------+--------+------------------+
+| Implant replacement              | £45,000         |        |                  |
++----------------------------------+-----------------+--------+------------------+
+| Implant complication             | £28,839         |        | @Smith2013       |
++----------------------------------+-----------------+--------+------------------+
+| Non-fatal shock                  | £22,880         |        | UK Stroke Assoc. |
++----------------------------------+-----------------+--------+------------------+
+| HCM without ICD                  | 0               |        |                  |
++----------------------------------+-----------------+--------+------------------+
+| SCD                              | 0               |        |                  |
++----------------------------------+-----------------+--------+------------------+
+| All-cause death                  | 0               |        |                  |
++----------------------------------+-----------------+--------+------------------+
 |                                  |                 |        |                  |
++----------------------------------+-----------------+--------+------------------+
+| *Probabilities*                  |                 |        |                  |
++----------------------------------+-----------------+--------+------------------+
+| Initial implant complication     | 0.047           |        | @Smith2013       |
++----------------------------------+-----------------+--------+------------------+
+| Replacement implant complication | 0.032           |        | @Smith2013       |
++----------------------------------+-----------------+--------+------------------+
+|                                  |                 |        |                  |
++----------------------------------+-----------------+--------+------------------+
+| Time horizon                     | 12 years        |        |                  |
++----------------------------------+-----------------+--------+------------------+
+| Implant replacement              | 10 years        |        |                  |
++----------------------------------+-----------------+--------+------------------+
+| Annual number of appointments    | 2               |        |                  |
++----------------------------------+-----------------+--------+------------------+
 
-: Model parameter values.
+: Model parameter values. $^*$either one-off/on state entry or recurring.
+
+Table 2 gives the starting state populations for non-zero states.
+
++-----------------+---------+------------+
+| Risk rule       | State   | Population |
++=================+=========+============+
+| Observed        | HCM ICD | 559        |
++-----------------+---------+------------+
+| \-              | HCM     | 3113       |
++-----------------+---------+------------+
+| Score \> 4%     | HCM ICD | 2561       |
++-----------------+---------+------------+
+| \-              | HCM     | 1111       |
++-----------------+---------+------------+
+| Score \> 6%     | HCM ICD | 542        |
++-----------------+---------+------------+
+| \-              | HCM     | 3130       |
++-----------------+---------+------------+
+| Risk factor \>0 | HCM ICD | 1785       |
++-----------------+---------+------------+
+| \-              | HCM     | 1887       |
++-----------------+---------+------------+
+| Risk factor \>1 | HCM ICD | 481        |
++-----------------+---------+------------+
+| \-              | HCM     | 3191       |
++-----------------+---------+------------+
+| Risk factor \>2 | HCM ICD | 78         |
++-----------------+---------+------------+
+| \-              | HCM     | 3594       |
++-----------------+---------+------------+
 
@@
 
-This gives us starting state populations for HCM with ICD and HCM without ICD which will be different for each risk decision rule.
+### Markov model
+
+Th patient data give us starting state populations for HCM with ICD and HCM without ICD which will be different for each risk decision rule.
 Further, the transition probabilities from these states will differ because of the case mixes.
@@
 
-We have used a year step size but if eg cost are accrued at different intervals then this can be adapted.
-The time horizon is set at 12 years but can be changed.
+We have used a year step size but if e.g. cost are accrued at different intervals then this can be adapted.
+The time horizon is set at 12 years from time of implant.
+
+### State health and cost equations
+
+We assume that an ICD patient has 2 annual appointments.
+All shocks are treated the same in terms of costs and health impact.
+Implantation can have complications.
+The cost of an implant complication is taken as a weighted sum of infection and dislodgement cost with values from [@Smith2013].
+
+-   Annual health: $$
+    e^0_{s=1}(t) = e^1_{s=1}(t) = u\_hcm + u\_icd
+    $$
+
+$$
+e^0_{s=2}(t) = e^1_{s=2}(t) = u\_shock
+$$
+
+$$
+e^0_{s}(t) = e^1_{s}(t) = 0, \;\; s = 3,4,5,6
+$$
+
+-   Initial costs: $$
+    c^0_{s=1}(t = 0) = c\_icd + p\_compl \times c\_compl
+    $$
+
+$$
+c^1_{s=1}(t = 0) = c\_icd + c\_rscore + p\_compl \times c\_compl
+$$
+
+-   Annual cost: $$
+    c^0_{s=1}(t) = c^1_{s=1}(t) = 2 c\_icd\_appt, \;\; t \neq t\_repl
+    $$
+
+$$
+c^0_{s=2}(t) = c^1_{s=2}(t) = c\_nfatal
+$$
+
+-   Implant replacement cost $$
+    c^0_{s=1}(t = t\_repl) = c^1_{s=1}(t = t\_repl) = 2 c\_icd\_appt + c\_icd\_repl + p\_compl \times c\_compl
+    $$
 
@@
 
-Using WinBUGS called from R, each new data set is used to generate posterior samples of transition probabilities.
+Using WinBUGS [@Zhang2014] called from R, each new data set is used to generate posterior samples of transition probabilities.
 Denote $x$ as the observed number of transitions, $p$ the probability of a transition and $n$ as the total number of transitions from a given state.
@@
 
-$$x^{(1)}_{i.} \sim \mbox{Multinomial}(p^{(1)}_{i.}, n^{(1)}_i), \;\; i = 1,3$$
-$$x^{(2)}_{i.} \sim \mbox{Multinomial}(p^{(2)}_{i.}, n^{(2)}_i), \;\; i = 1,3$$
+$$x^{(1)}_{i.} \sim \mbox{Multinomial}(p^{(1)}_{i.}, n^{(1)}_i), \;\; i = 1,3$$ $$x^{(2)}_{i.} \sim \mbox{Multinomial}(p^{(2)}_{i.}, n^{(2)}_i), \;\; i = 1,3$$
 
-$$p^{(1)}_{i.} \sim \mbox{Dirichlet}(\boldsymbol{\alpha}^{(1)} ), \;\; i = 1,3$$
-$$p^{(2)}_{i.} \sim \mbox{Dirichlet}(\boldsymbol{\alpha}^{(2)} ), \;\; i = 1,3$$
+$$p^{(1)}_{i.} \sim \mbox{Dirichlet}(\boldsymbol{\alpha}^{(1)} ), \;\; i = 1,3$$ $$p^{(2)}_{i.} \sim \mbox{Dirichlet}(\boldsymbol{\alpha}^{(2)} ), \;\; i = 1,3$$
 

### 2021-04-20 HCM.SCD.CEanalysis/R/ce_sim.R

@@ -45,6 +45,10 @@ ce_sim <- function(pop,
 
   for (k in seq_len(n_interv)) {
     for (i in seq_len(n_sim)) {
+
+      # sample costs
+      rc_unit <- c_unit()
+
       for (j in seq_len(tmax)) {
 
         # time-homogeneous dim 1
@@ -62,19 +66,22 @@ ce_sim <- function(pop,
 
         e_state <- e_unit[[k]]*(1 - pdecr[[k]])^j
 
-        cost[[k]][i, j] <- c_unit[[j]][[k]] %*% pop[[k]][, j, i]
+        cost[[k]][i, j] <- rc_unit[[j]][[k]] %*% pop[[k]][, j, i]
 
         eff[[k]][i, j] <- e_state %*% pop[[k]][, j, i]
 
         dcost[[k]][i, j] <- cost[[k]][i, j] / disc
         deff[[k]][i, j] <- eff[[k]][i, j] / disc
       }
-    }
 
-    if (!any(is.na(c_init))) {
       # add one-off starting state costs
-      cost[[k]][, 1] <- dcost[[k]][, 1] <-
-        c_init[[k]] %*% pop[[k]][,1,1]
+      if (is.function(c_init)) {
+        cost[[k]][i, 1] <-
+          cost[[k]][i, 1] + c_init()[[k]] %*% pop[[k]][,1,1]
+
+        dcost[[k]][i, 1] <-
+          dcost[[k]][i, 1] + c_init()[[k]] %*% pop[[k]][,1,1]
+      }
     }
   }
 
@@ -83,5 +90,13 @@ ce_sim <- function(pop,
        dcost = dcost,
        eff = eff,
        deff = deff)
+}
+
+
+#
+next_pop <- function(pop, k, i, j, t) {
+  function(s) {
+    t(pop[[k]][, j - 1, i]) %*% probs[[k]][, s, t, i]
+  }
 }
 

### 2021-04-20 HCM.SCD.CEanalysis/scripts/main_ce_analysis.R

@@ -23,6 +23,11 @@ data("ipd_risk")
 # 5. SCD
 # 6. all-cause mortality
 
+# structure:
+# (1.icd  2.shock 3.mortality)
+# (4.~icd 5.scd   6.mortality)
+
+
 # start state populations
 init_risk6 <- table(ipd_risk$risk_over_6)
 init_obs <- table(ipd_risk$icd)
@@ -37,49 +42,82 @@ n_sim <- nrow(lambda.0)
 S <- length(n_init$obs)
 J <- 12  # max time (year)
 
+
 # utilities and costs --------
 # for each intervention
 
-# managed with an ICD (Noyes 2007)
-u_hcm <- 0.637
-u_icd <- -0.05
-u_shock <- 0.5
+u_hcm <- 0.637   # managed with an ICD (Noyes 2007)
+u_icd <- -0.05   # ICD annual detriment
+u_shock <- 0.5   # annual detriment
+u_icd_init <- -6/365 # icd implant one-off (Smith EJH 2013)
 
 e_unit <- list(obs =   c(u_hcm + u_icd, u_shock, 0,
                          u_hcm,         0,       0),
                risk6 = c(u_hcm + u_icd, u_shock, 0,
                          u_hcm,         0,       0))
 
-# cost of non-fatal HCM related events £22,880 (UK Stroke Association)
-c_nfatal <- 22880
 
-# EY02B NHS tariffs implantation cost
-c_icd <- 4666
+# non-fatal HCM related events (UK Stroke Association)
+c_nfatal <- 22880
+c_icd <- 4666         # EY02B NHS tariffs implantation cost
 c_rscore <- 20
 c_icd_appt <- 10
 c_icd_repl <- 45000
 
-c_init <- list(obs =   c(c_icd, 0, 0,
-                         0,     0, 0),
-               risk6 = c(c_icd + c_rscore, 0, 0,
-                         0,                0, 0))
+# implant complication
+p_compl <- 0.047        # Smith EHJ 2013
+p_compl_repl <- 0.032   # Smith EHJ 2013
+# weighted infection, dislodgement
+c_compl <- (0.0227*37116 + 0.00828*6146)/(0.0227 + 0.00828)
 
 c_entry <- list(obs =   c(0, 0, 0,
                           0, 0, 0),
                 risk6 = c(0, 0, 0,
                           0, 0, 0))
 
[…]

### 2021-04-20 HCM.SCD.CEanalysis/tests/testthat/test-ce_sim.R

@@ -1,0 +1,48 @@
+
+
+S <- 2
+n_sim <- 2
+p <- array(c(1,0,
+             0,1,
+             0,1,
+             1,0),
+           dim = c(S, S, 1, n_sim))
+t_lim <- 10
+
+# interv
+probs = list(p, p)
+
+# time > interv > state
+c_unit <- function(t_lim) {
+  function()
+    purrr::map(1:t_lim,
+               ~list(c(1,1), c(1,1)))
+}
+
+# interv > state
+c_init <- function()
+  list(c(1,1), c(1,1))
+
+# interv > state
+e_unit <- list(c(1,1), c(0,0))
+
+# interv > time > state
+pop <- list(array(c(1,0,NA,NA),
+                  dim = c(S, t_lim, 2)),
+            array(c(1,0,NA,NA),
+                  dim = c(S, t_lim, 2)))
+
+test_that("basic inputs", {
+
+  ce_sim(pop = pop,
+         probs = probs,
+         c_unit = c_unit(t_lim),
+         e_unit = e_unit,
+         c_init = c_init)
+})
+
+
+
+
+
+

### 2021-04-21 HCM.SCD.CEanalysis/R/ce_sim_refac.R
pulled-out ce_markov() for single sim and intervention to test * todo: integrate back in to ce_sim()

@@ -1,0 +1,109 @@
+
+##TODO...
+
+# ce_sim <- function(pop,
+#                    probs,
+#                    c_unit,
+#                    e_unit,
+#                    c_init = NA,
+#                    pdecr = NA,
+#                    delta = 0.035) {
+#
+#   stopifnot(delta >= 0, delta <= 1)
+#
+#   ##TODO: check inputs
+#
+#   ##TODO: interleave pop, probs, c_unit, c_init. e_unit
+#   # so that list over interventions and sim i
+#   # inputs <-
+#
+#   out <-
+#     map(inputs,
+#         ~ map(.x, ~ce_markov))
+#
+#   # reorder
+#   # purrr::
+#
+#   list(pop = pop,
+#        cost = cost,
+#        dcost = dcost,
+#        eff = eff,
+#        deff = deff)
+# }
+#
+#
+# #
+# next_pop <- function(pop, j, t) {
+#   function(s) {
+#     t(pop[, j - 1]) %*% probs[, s, t]
+#   }
+# }
+
+
+#
+ce_markov <- function(inputs,
+                      pdecr = NA,
+                      delta = 0.035) {
+
+  pop <- inputs$pop
+  probs <- inputs$probs
+  c_unit <- inputs$c_unit
+  c_init <- inputs$c_init
+  e_unit <- inputs$e_unit
+
+  S <- dim(pop)[1]        # number of states
+  tmax <- dim(pop)[2]     # time horizon
+
+  if (any(is.na(pdecr))) {
+    pdecr <- rep(0, S)
+  }
+
+  # sample costs
+  rc_unit <- c_unit()
+
+  cost <- vector("numeric", tmax)
+  eff <- vector("numeric", tmax)
+  dcost <- vector("numeric", tmax)
+  deff <- vector("numeric", tmax)
+
+  for (j in seq_len(tmax)) {
+
+    # time-homogeneous dim 1
+    t <- min(dim(probs)[3], j - 1)
+
+    if (j > 1) {
+      for (s in seq_len(S)) {
+
+        pop[s, j] <-
+          t(pop[, j - 1]) %*% probs[, s, t]
+      }
+    }
+
+    disc <- (1 + delta)^(j-1)
+
+    e_state <- e_unit*(1 - pdecr)^j
+
+    cost[j] <- rc_unit[[j]] %*% pop[, j]
+
+    eff[j] <- e_state %*% pop[, j]
+
[…]

### 2021-04-21 HCM.SCD.CEanalysis/tests/testthat/test-ce_markov.R

@@ -1,0 +1,44 @@
+
+library(dplyr)
+library(purrr)
+
+
+pop_init <- c(1,0)
+
+pop <- array(pop_init,
+             dim = c(S, t_lim))
+
+S <- length(pop_init)
+
+probs <- array(c(0,1,
+                 1,0),
+               dim = c(S, S, 1))
+t_lim <- 10
+
+# time=dependent
+c_unit <- function(t_lim) {
+  function()
+    purrr::map(1:t_lim, ~ c(1,1))
+}
+
+c_init <- function() c(1,1)
+
+# interv > state
+e_unit <- c(1,0)
+
+
+list(pop = pop,
+     probs = probs,
+     c_unit = c_unit(t_lim),
+     c_init = c_init,
+     e_unit = e_unit) %>%
+ce_markov()
+
+test_that("basic inputs", {
+  # expect_equal()
+})
+
+
+
+
+

### 2021-07-21 HCM.SCD.CEanalysis/inst/rmarkdown/draft.Rmd
added all the components for the paper but the results are not final; they need to be run properly

@@
 title: "HCM SCD risk algorithm cost-effectiveness analysis"
-author: "Nathan Green"
+author: "Nathan Green, Yang Chen"
 date: "09/11/2020"
-output: pdf_document
+output: 
+  bookdown::pdf_document2:
+    toc: false
+    keep_tex: yes
 editor_options: 
@@
     wrap: sentence
+params:
+  data: "ipd_risk"
 bibliography: bibliography.bib
@@
 
+```{r, echo=FALSE}
+data(list = params$data, package = "HCM.SCD.CEanalysis")
+```
+
+```{r echo=FALSE}
+centre_tab <- dplyr::count(ipd_risk, centre)
+centre <- centre_tab$centre
+n <- centre_tab$n
+
+cens_tab <- dplyr::count(ipd_risk, d)
+cens <- cens_tab$d
+nd <- cens_tab$n
+```
+
 ## Introduction
@@
 
-There have been previous cost-effectiveness analyses [@Magnusson2020; @Boriani2014; @Bryant2007; @Caro2007; @Cowie2009; @Garcia-Perez2015; @Mealing2016; @Sanders2005; @Smith2013]
+There have been previous cost-effectiveness analyses of ICD implants [@Magnusson2020; @Boriani2014; @Bryant2007; @Caro2007; @Cowie2009; @Garcia-Perez2015; @Mealing2016; @Sanders2005; @Smith2013]
+
+Cost-effectiveness of a cardiovascular risk prediction algorithm.
+[@Zomer2017] for statins interventions and people with severe mental illness.
+
+[@Yao2007] use a Markov model and consider multiple implantations attempts if unsuccessful.
+[@Colquitt2014] is a Health Technology Assessment which compares optimal pharmacological therapy (OPT) with or without ICD.
+[@Tomini2016] is a review of economic evaluation models for cardiac resynchronization therapy with implantable cardioverter defibrillators in patients with heart failure.
+[@Ommen2020] 2020 AHA/ACC Guideline for the Diagnosis and Treatment of Patients With Hypertrophic Cardiomyopathy.
 
@@
 The main data set contains individual-level follow-up data of patients with HCM who may have been given an ICD due to some risk decision.
+Key cohort characteristics include the following.
+Patients were enrolled from the 6 health centres: `r paste0(centre[1], " (", n[1], ")")`, `r paste0(centre[2], " (", n[2], ")")`, `r paste0(centre[3], " (", n[3], ")")`, `r paste0(centre[4], " (", n[4], ")")`, `r paste0(centre[5], " (", n[5], ")")`, `r paste0(centre[6], " (", n[6], ")")`; The amount of censoring was `r paste0(cens[1], " (", nd[1], ")")`, `r paste0(cens[2], " (", nd[2], ")")`; The mean age (sd) was `r round(mean(ipd_risk$age),0)` (`r round(sd(ipd_risk$age),0)`); The start of study data collection was in `r min(ipd_risk$year)` to `r max(ipd_risk$year)`.
+Further plots are given in the Appendix.
+See [@OMahony2014] for further details.
 
@@
 
-+----------------------------------+-----------------+--------+------------------+
-| Description                      | Value$^*$       | Range  | Source           |
-+==================================+=================+========+==================+
-| *Health*                         |                 |        |                  |
-+----------------------------------+-----------------+--------+------------------+
-| Manage with ICD                  | 0.637 QALY/year |        | @Noyes2007       |
-+----------------------------------+-----------------+--------+------------------+
-| Implantation procedure utility   | -0.016          |        | @Smith2013       |
-+----------------------------------+-----------------+--------+------------------+
-| Shock utility                    | -0.5            |        |                  |
-+----------------------------------+-----------------+--------+------------------+
-| HCM without ICD                  | 1 QALY/year     |        |                  |
-+----------------------------------+-----------------+--------+------------------+
-| Death                            | 0 QALY/year     |        |                  |
-+----------------------------------+-----------------+--------+------------------+
-|                                  |                 |        |                  |
-+----------------------------------+-----------------+--------+------------------+
-| *Cost*                           |                 |        |                  |
-+----------------------------------+-----------------+--------+------------------+
-| ICD appointment                  | £10             |        |                  |
-+----------------------------------+-----------------+--------+------------------+
-| Perform risk score               | £20?            |        |                  |
-+----------------------------------+-----------------+--------+------------------+
-| Implant ICD                      | £4,666          |        | EY02B Tariffs    |
-+----------------------------------+-----------------+--------+------------------+
-| Implant replacement              | £45,000         |        |                  |
-+----------------------------------+-----------------+--------+------------------+
-| Implant complication             | £28,839         |        | @Smith2013       |
-+----------------------------------+-----------------+--------+------------------+
-| Non-fatal shock                  | £22,880         |        | UK Stroke Assoc. |
-+----------------------------------+-----------------+--------+------------------+
-| HCM without ICD                  | 0               |        |                  |
-+----------------------------------+-----------------+--------+------------------+
-| SCD                              | 0               |        |                  |
-+----------------------------------+-----------------+--------+------------------+
-| All-cause death                  | 0               |        |                  |
-+----------------------------------+-----------------+--------+------------------+
-|                                  |                 |        |                  |
-+----------------------------------+-----------------+--------+------------------+
-| *Probabilities*                  |                 |        |                  |
-+----------------------------------+-----------------+--------+------------------+
-| Initial implant complication     | 0.047           |        | @Smith2013       |
-+----------------------------------+-----------------+--------+------------------+
-| Replacement implant complication | 0.032           |        | @Smith2013       |
-+----------------------------------+-----------------+--------+------------------+
-|                                  |                 |        |                  |
-+----------------------------------+-----------------+--------+------------------+
-| Time horizon                     | 12 years        |        |                  |
-+----------------------------------+-----------------+--------+------------------+
-| Implant replacement              | 10 years        |        |                  |
-+----------------------------------+-----------------+--------+------------------+
-| Annual number of appointments    | 2               |        |                  |
-+----------------------------------+-----------------+--------+------------------+
++----------------------------------+-------------+-----------------+-------+------------------+
+| Description                      | Parameter   | Value$^*$       | Range | Source           |
++==================================+=============+=================+=======+==================+
+| *Health*                         |             |                 |       |                  |
++----------------------------------+-------------+-----------------+-------+------------------+
+| Manage with ICD                  | `u_icd`     | 0.637 QALY/year |       | @Noyes2007       |
++----------------------------------+-------------+-----------------+-------+------------------+
+| Implantation procedure utility   | `u_implant` | -0.016          |       | @Smith2013       |
++----------------------------------+-------------+-----------------+-------+------------------+
+| Shock utility                    | `u_shock`   | -0.5            |       |                  |
++----------------------------------+-------------+-----------------+-------+------------------+
+| HCM without ICD                  | `u_hcm`     | 1 QALY/year     |       |                  |
++----------------------------------+-------------+-----------------+-------+------------------+
+| Death                            | `u_death`   | 0 QALY/year     |       |                  |
++----------------------------------+-------------+-----------------+-------+------------------+
+|                                  |             |                 |       |                  |
++----------------------------------+-------------+-----------------+-------+------------------+
+| *Cost*                           |             |                 |       |                  |
++----------------------------------+-------------+-----------------+-------+------------------+
+| ICD appointment                  | `c_appt`    | £10             |       |                  |
++----------------------------------+-------------+-----------------+-------+------------------+
+| Perform risk score               | `c_rs`      | £20?            |       |                  |
++----------------------------------+-------------+-----------------+-------+------------------+
+| Implant ICD                      | `c_implant` | £4,666          |       | EY02B Tariffs    |
++----------------------------------+-------------+-----------------+-------+------------------+
+| Implant replacement              | `c_repl`    | £45,000         |       |                  |
++----------------------------------+-------------+-----------------+-------+------------------+
+| Implant complication             | `c_compl`   | £28,839         |       | @Smith2013       |
++----------------------------------+-------------+-----------------+-------+------------------+
+| Non-fatal shock                  | `c_shock`   | £22,880         |       | UK Stroke Assoc. |
++----------------------------------+-------------+-----------------+-------+------------------+
+| HCM without ICD                  | `c_hcm`     | 0               |       |                  |
++----------------------------------+-------------+-----------------+-------+------------------+
+| SCD                              | `c_scd`     | 0               |       |                  |
++----------------------------------+-------------+-----------------+-------+------------------+
+| All-cause death                  | `c_death`   | 0               |       |                  |
++----------------------------------+-------------+-----------------+-------+------------------+
+|                                  |             |                 |       |                  |
++----------------------------------+-------------+-----------------+-------+------------------+
+| *Probabilities*                  |             |                 |       |                  |
++----------------------------------+-------------+-----------------+-------+------------------+
+| Initial implant complication     | `p_compl`   | 0.047           |       | @Smith2013       |
++----------------------------------+-------------+-----------------+-------+------------------+
+| Replacement implant complication | `p_repl`    | 0.032           |       | @Smith2013       |
++----------------------------------+-------------+-----------------+-------+------------------+
+|                                  |             |                 |       |                  |
++----------------------------------+-------------+-----------------+-------+------------------+
+| Time horizon                     | `T`         | 12 years        |       |                  |
++----------------------------------+-------------+-----------------+-------+------------------+
+| Implant replacement              | `t_repl`    | 10 years        |       |                  |
++----------------------------------+-------------+-----------------+-------+------------------+
+| Annual number of appointments    | n_appt      | 2               |       |                  |
++----------------------------------+-------------+-----------------+-------+------------------+
 
@@
 
+: Starting state populations by decision rule.
+
 ## Methods
@@
 -   ICD given if number of risk factors \> 0, 1 or 2
--   ICD given if risk score \>6%
--   ICD given if risk score \>4%
+-   ICD given if risk score \> 6%
+-   ICD given if risk score \> 4%
 
@@
 
-A diagram of the current cohort model is given below.
+A diagram of the current cohort model is given in Figure \@ref(fig:model).
 
-```{r fig.align="center", echo=FALSE}
-knitr::include_graphics("model_diagram.png")
+```{r model, fig.align="center", echo=FALSE, fig.cap="Markov model diagram. Bold circles represent starting states and dashed circles represent sink states."}
+knitr::include_graphics("../../images/model_diagram.png")
 ```
 
-Assuming that shocked patients return to the HCM ICD state then the transition matrix may look like the following.
+Assuming that shocked patients return to the HCM ICD state then the transition matrix looks like the following.
 
@@
 The cost of an implant complication is taken as a weighted sum of infection and dislodgement cost with values from [@Smith2013].
+Subscript $s$ denotes the state number and superscript denotes the intervention either ICD or not.
 
 -   Annual health: $$
-    e^0_{s=1}(t) = e^1_{s=1}(t) = u\_hcm + u\_icd
+    e^0_{s=1}(t) = e^1_{s=1}(t) = \mbox{u\_hcm} + \mbox{u\_icd}
+    $$ $$
+    e^0_{s=3}(t) = e^1_{s=3}(t) = \mbox{u\_hcm}
+    $$ $$
+    e^0_{s=2}(t) = e^1_{s=2}(t) = \mbox{u\_shock}
     $$
@@
 $$
-e^0_{s=2}(t) = e^1_{s=2}(t) = u\_shock
+e^0_{s}(t) = e^1_{s}(t) = 0, \;\; s = 4,5,6
+$$
+
+-   Initial costs: $$
+    c^0_{s=1}(t = 0) = \mbox{c\_icd} + \mbox{p\_compl} \times \mbox{c\_compl}
     $$
@@
 $$
-e^0_{s}(t) = e^1_{s}(t) = 0, \;\; s = 3,4,5,6
+c^1_{s=1}(t = 0) = \mbox{c\_icd} + \mbox{c\_rscore} + \mbox{p\_compl} \times \mbox{c\_compl}
 $$
 
--   Initial costs: $$
-    c^0_{s=1}(t = 0) = c\_icd + p\_compl \times c\_compl
+-   Annual cost: $$
+    c^0_{s=1}(t) = c^1_{s=1}(t) = 2 \mbox{c\_appt}, \;\; t \neq \mbox{t\_repl}
     $$
@@
 $$
-c^1_{s=1}(t = 0) = c\_icd + c\_rscore + p\_compl \times c\_compl
-$$
-
--   Annual cost: $$
-    c^0_{s=1}(t) = c^1_{s=1}(t) = 2 c\_icd\_appt, \;\; t \neq t\_repl
-    $$
-
-$$
-c^0_{s=2}(t) = c^1_{s=2}(t) = c\_nfatal
+c^0_{s=2}(t) = c^1_{s=2}(t) = \mbox{c\_shock}
 $$
@@
 -   Implant replacement cost $$
-    c^0_{s=1}(t = t\_repl) = c^1_{s=1}(t = t\_repl) = 2 c\_icd\_appt + c\_icd\_repl + p\_compl \times c\_compl
+    c^0_{s=1}(t) = c^1_{s=1}(t) = 2 \mbox{c\_appt} + \mbox{c\_repl} + \mbox{p\_compl} \times \mbox{c\_compl}, \;\;  t = \mbox{t\_repl}
     $$
@@
 
-$$x^{(1)}_{i.} \sim \mbox{Multinomial}(p^{(1)}_{i.}, n^{(1)}_i), \;\; i = 1,3$$ $$x^{(2)}_{i.} \sim \mbox{Multinomial}(p^{(2)}_{i.}, n^{(2)}_i), \;\; i = 1,3$$
+$$x^{(1)}_{i.} \sim \mbox{Multinomial}(p^{(1)}_{i.}, n^{(1)}_i), \;\; i = 1,3$$
 
-$$p^{(1)}_{i.} \sim \mbox{Dirichlet}(\boldsymbol{\alpha}^{(1)} ), \;\; i = 1,3$$ $$p^{(2)}_{i.} \sim \mbox{Dirichlet}(\boldsymbol{\alpha}^{(2)} ), \;\; i = 1,3$$
+$$x^{(2)}_{i.} \sim \mbox{Multinomial}(p^{(2)}_{i.}, n^{(2)}_i), \;\; i = 1,3$$
+
+$$p^{(1)}_{i.} \sim \mbox{Dirichlet}(\boldsymbol{\alpha}^{(1)} ), \;\; i = 1,3$$
+
+$$p^{(2)}_{i.} \sim \mbox{Dirichlet}(\boldsymbol{\alpha}^{(2)} ), \;\; i = 1,3$$
 
@@
 
-## Workflow
-
-The main files to perform the analysis are:
-
--   `prep_study_data.R` munges the raw data and save the input data in `data/`
--   `BUGS/script.R` runs the BUGS code in `BUGS/model.txt`
--   `main-ce-analysis.R` performs the cost-effectiveness analysis
--   `pop_counts_plot.R` creates output plots
-
 ## Results
 
-We give some example output to demonstrate what will be produced with the final model.
+We give results of the model fitting and cost-effectiveness analysis.
 
-Histogram of posterior distributions for state transition probabilities.
+### Model fitting
 
-![](posterior_histograms.png)
+Figure \@ref(fig:hist) gives a histogram of posterior distributions for state transition probabilities.
 
-Below is an example of state occupancy over time plot.
-This shows that for the new algorithm there are fewer scd and more shocks.
+```{r hist, fig.align="center", echo=FALSE, out.width = "80%", fig.cap="Histogram of posterior distributions for state transition probabilities."}
+knitr::include_graphics("posterior_histograms.png")
+```
 
-![](../../images/state_pop_over_time.png)
+Figure \@ref(fig:statepop) gives an example of state occupancy over time plot.
+This shows that for the new algorithm there are fewer SCD and more shocks.
+
+```{r statepop, fig.align="center", echo=FALSE, fig.cap="State occupancy over time."}
+knitr::include_graphics("../../images/state_pop_over_time.png")
+```
+
+### Cost-effectiveness
+
+| Strategy          | Cost, $c$ | $\Delta c$ | QALYs, $e$ | $\Delta e$ | ICER |
+|-------------------|-----------|------------|------------|------------|------|
+| Baseline          |           |            |            |            |      |
+| \> 4%             |           |            |            |            |      |
+| \> 6%             |           |            |            |            |      |
+| \> 1 risk factor  |           |            |            |            |      |
+| \> 2 risk factors |           |            |            |            |      |
+
+: Cost-effectiveness statistics
+
+Figure \@ref(fig:ceac) gives the CEAC and 
+Figure \@ref(fig:ceplane) gives the CE-plane.
+
+```{r ceac, fig.align="center", echo=FALSE, fig.cap="Cost-effectiveness acceptability curve (CEAC)."}
+knitr::include_graphics("../../images/ceac.png")
+```
+
+```{r ceplane, fig.align="center", echo=FALSE, fig.cap="Cost-effectiveness plane."}
+knitr::include_graphics("../../images/ceplane.png")
+```
+
+## Discussion
+
+<!-- shared decision-making -->
+
+Decision making as to whether have the implant is shared between patient and medical professional.
+The SCD risk score is a support tool which contributes to the final decision.
+An SCD risk score of 6% is treatment recommendation for one patient versus another.
+There could be individual preferences that out-weighs the risk score and mean a patient chooses the alternative.
+
+<!-- better QoL -->
+
+Having more robust quality of life data will help to make such decisions.
+more embedded PROMs research
+
+Informatics Consult [@Lai2021] automation to scale evidence generation and to accelerate the return of results within clinical time-scales.
+
+<!-- limitation -->
+
+We did not include CRT-D or s-ICD but that can be overcome by sensitivity analyses that show e.g.
+the main message is that conclusions about the ICER etc will be robust to most departures from base assumptions other than tweaks in utility (which are the part of the model based on the weakest data!)
+
+[@Olivotto2020] The role of new DMARDs and their ability to influence quality of life.
+So if both ICD and non-ICD patients experience significant increases in QoL, are these equally distributed?
+
+# Appendix
+
+```{r barplot, fig.align="center", out.width = "80%", echo=FALSE, fig.cap="Year of study entry bar plot."}
+knitr::include_graphics("../../images/year_of_entry_barplot.png")
+```
 

### 2021-07-21 HCM.SCD.CEanalysis/inst/rmarkdown/draft.tex
added all the components for the paper but the results are not final; they need to be run properly

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={HCM SCD risk algorithm cost-effectiveness analysis},
  pdfauthor={Nathan Green, Yang Chen},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{HCM SCD risk algorithm cost-effectiveness analysis}
\author{Nathan Green, Yang Chen}
\date{09/11/2020}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Sudden cardiac death (SCD) risk score algorithm details see (O'Mahony et al. 2014).

There have been previous cost-effectiveness analyses of ICD implants (Magnusson and Wimo 2020; Boriani et al. 2014; Bryant et al. 2007; Caro et al. 2007; Cowie et al. 2009; García-Pérez et al. 2015; Mealing et al. 2016; Sanders, Hlatky, and Owens 2005; Smith et al. 2013)

Cost-effectiveness of a cardiovascular risk prediction algorithm.
(Zomer et al. 2017) for statins interventions and people with severe mental illness.

(Yao et al. 2007) use a Markov model and consider multiple implantations attempts if unsuccessful.
(Colquitt et al. 2014) is a Health Technology Assessment which compares optimal pharmacological therapy (OPT) with or without ICD.
(Tomini 2016) is a review of economic evaluation models for cardiac resynchronization therapy with implantable cardioverter defibrillators in patients with heart failure.
(Ommen et al. 2020) 2020 AHA/ACC Guideline for the Diagnosis and Treatment of Patients With Hypertrophic Cardiomyopathy.

\hypertarget{data}{%
\subsection{Data}\label{data}}

The main data set contains individual-level follow-up data of patients with HCM who may have been given an ICD due to some risk decision.
Key cohort characteristics include the following.
Patients were enrolled from the 6 health centres: Athens (474), Bologna (456), Coruna (590), London (1592), Murcia (404), Naples (156); The amount of censoring was 0 (3475), 1 (197); The mean age (sd) was 48 (16); The start of study data collection was in 1972 to 2011. Further plots are given in the Appendix. See (O'Mahony et al. 2014) for further details.

Health and cost data were obtained from literature and expert opinion.
What values we wish to use will determine the form of the state model.
For instance, if cost are only accrued on entry to a state then we may use a tunnel state.
If cost depend on the patient history then we may need to duplicate a state.

Table 1 gives the unit cost and health values used in the model.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.37}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.15}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.19}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.09}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.20}}@{}}
\caption{Model parameter values. \(^*\)either one-off/on state entry or recurring.}\tabularnewline
\toprule
Description & Parameter & Value\(^*\) & Range & Source \\
\midrule
\endfirsthead
\toprule
Description & Parameter & Value\(^*\) & Range & Source \\
\midrule
\endhead
\emph{Health} & & & & \\
Manage with ICD & \texttt{u\_icd} & 0.637 QALY/year & & Noyes et al. (2007) \\
Implantation procedure utility & \texttt{u\_implant} & -0.016 & & Smith et al. (2013) \\
Shock utility & \texttt{u\_shock} & -0.5 & & \\
HCM without ICD & \texttt{u\_hcm} & 1 QALY/year & & \\
Death & \texttt{u\_death} & 0 QALY/year & & \\
& & & & \\
\emph{Cost} & & & & \\
ICD appointment & \texttt{c\_appt} & £10 & & \\
Perform risk score & \texttt{c\_rs} & £20? & & \\
Implant ICD & \texttt{c\_implant} & £4,666 & & EY02B Tariffs \\
Implant replacement & \texttt{c\_repl} & £45,000 & & \\
Implant complication & \texttt{c\_compl} & £28,839 & & Smith et al. (2013) \\
Non-fatal shock & \texttt{c\_shock} & £22,880 & & UK Stroke Assoc. \\
HCM without ICD & \texttt{c\_hcm} & 0 & & \\
SCD & \texttt{c\_scd} & 0 & & \\
All-cause death & \texttt{c\_death} & 0 & & \\
& & & & \\
\emph{Probabilities} & & & & \\
Initial implant complication & \texttt{p\_compl} & 0.047 & & Smith et al. (2013) \\
Replacement implant complication & \texttt{p\_repl} & 0.032 & & Smith et al. (2013) \\
& & & & \\
Time horizon & \texttt{T} & 12 years & & \\
Implant replacement & \texttt{t\_repl} & 10 years & & \\
Annual number of appointments & n\_appt & 2 & & \\
\bottomrule
\end{longtable}

Table 2 gives the starting state populations for non-zero states.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.25}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.14}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.18}}@{}}
\caption{Starting state populations by decision rule.}\tabularnewline
\toprule
Risk rule & State & Population \\
\midrule
\endfirsthead
\toprule
Risk rule & State & Population \\
\midrule
\endhead
Observed & HCM ICD & 559 \\
- & HCM & 3113 \\
Score \textgreater{} 4\% & HCM ICD & 2561 \\
- & HCM & 1111 \\
Score \textgreater{} 6\% & HCM ICD & 542 \\
- & HCM & 3130 \\
Risk factor \textgreater0 & HCM ICD & 1785 \\
- & HCM & 1887 \\
Risk factor \textgreater1 & HCM ICD & 481 \\
- & HCM & 3191 \\
Risk factor \textgreater2 & HCM ICD & 78 \\
- & HCM & 3594 \\
\bottomrule
\end{longtable}

\hypertarget{methods}{%
\subsection{Methods}\label{methods}}

The individual-level patient data are first stratified in to two groups for each risk algorithm.
These are

\begin{itemize}
\tightlist
\item
  Partition observed in data set
\item
  ICD given if number of risk factors \textgreater{} 0, 1 or 2
\item
  ICD given if risk score \textgreater{} 6\%
\item
  ICD given if risk score \textgreater{} 4\%
\end{itemize}

We included the option of a fuzzy decision boundary such that near the threshold there is some random variation as to whether a patient received an ICD or not.

\hypertarget{markov-model}{%
\subsubsection{Markov model}\label{markov-model}}

Th patient data give us starting state populations for HCM with ICD and HCM without ICD which will be different for each risk decision rule.
Further, the transition probabilities from these states will differ because of the case mixes.

A diagram of the current cohort model is given in Figure \ref{fig:model}.

\begin{figure}

{\centering \includegraphics[width=3.25in]{../../images/model_diagram}

}

\caption{Markov model diagram. Bold circles represent starting states and dashed circles represent sink states.}\label{fig:model}
\end{figure}

Assuming that shocked patients return to the HCM ICD state then the transition matrix looks like the following.

\[
\begin{pmatrix}
p_{11} & p_{12} & 0 & 0 & p_{15}\\
1 & 0 & 0 & 0 & 0\\
0 & 0 & p_{33} & p_{34} & p_{35}\\
0 & 0 & 0 & 1 & 0\\
0 & 0 & 0 & 0 & 1
\end{pmatrix}
\]

We have used a year step size but if e.g.~cost are accrued at different intervals then this can be adapted.
The time horizon is set at 12 years from time of implant.

\hypertarget{state-health-and-cost-equations}{%
\subsubsection{State health and cost equations}\label{state-health-and-cost-equations}}

We assume that an ICD patient has 2 annual appointments.
All shocks are treated the same in terms of costs and health impact.
Implantation can have complications.
The cost of an implant complication is taken as a weighted sum of infection and dislodgement cost with values from (Smith et al. 2013).
Subscript \(s\) denotes the state number and superscript denotes the intervention either ICD or not.

\begin{itemize}
\tightlist
\item
  Annual health: \[
  e^0_{s=1}(t) = e^1_{s=1}(t) = \mbox{u\_hcm} + \mbox{u\_icd}
  \] \[
  e^0_{s=3}(t) = e^1_{s=3}(t) = \mbox{u\_hcm}
  \] \[
  e^0_{s=2}(t) = e^1_{s=2}(t) = \mbox{u\_shock}
  \]
\end{itemize}

\[
e^0_{s}(t) = e^1_{s}(t) = 0, \;\; s = 4,5,6
\]

\begin{itemize}
\tightlist
\item
  Initial costs: \[
  c^0_{s=1}(t = 0) = \mbox{c\_icd} + \mbox{p\_compl} \times \mbox{c\_compl}
  \]
\end{itemize}

\[
c^1_{s=1}(t = 0) = \mbox{c\_icd} + \mbox{c\_rscore} + \mbox{p\_compl} \times \mbox{c\_compl}
\]

\begin{itemize}
\tightlist
\item
  Annual cost: \[
  c^0_{s=1}(t) = c^1_{s=1}(t) = 2 \mbox{c\_appt}, \;\; t \neq \mbox{t\_repl}
  \]
\end{itemize}

\[
c^0_{s=2}(t) = c^1_{s=2}(t) = \mbox{c\_shock}
\]

\begin{itemize}
\tightlist
\item
  Implant replacement cost \[
  c^0_{s=1}(t) = c^1_{s=1}(t) = 2 \mbox{c\_appt} + \mbox{c\_repl} + \mbox{p\_compl} \times \mbox{c\_compl}, \;\;  t = \mbox{t\_repl}
  \]
\end{itemize}

\hypertarget{transition-probability-inference}{%
\paragraph{Transition probability inference}\label{transition-probability-inference}}

Using WinBUGS (Zhang 2014) called from R, each new data set is used to generate posterior samples of transition probabilities.
Denote \(x\) as the observed number of transitions, \(p\) the probability of a transition and \(n\) as the total number of transitions from a given state.
The hyperparameters \(\alpha\) characterise the prior knowledge on \(p\).
Superscripts indicate the decision rule used.

\[x^{(1)}_{i.} \sim \mbox{Multinomial}(p^{(1)}_{i.}, n^{(1)}_i), \;\; i = 1,3\]

\[x^{(2)}_{i.} \sim \mbox{Multinomial}(p^{(2)}_{i.}, n^{(2)}_i), \;\; i = 1,3\]

\[p^{(1)}_{i.} \sim \mbox{Dirichlet}(\boldsymbol{\alpha}^{(1)} ), \;\; i = 1,3\]

\[p^{(2)}_{i.} \sim \mbox{Dirichlet}(\boldsymbol{\alpha}^{(2)} ), \;\; i = 1,3\]

For all sink states,

\[
p^{(s)}_{ij} = \left\{
\begin{array}{ll}
1 & \mbox{if $i = j$};\\
0 & \mbox{if $i \neq j$}.
\end{array} \right.
\]

\hypertarget{results}{%
\subsection{Results}\label{results}}

We give results of the model fitting and cost-effectiveness analysis.

\hypertarget{model-fitting}{%
\subsubsection{Model fitting}\label{model-fitting}}

Figure \ref{fig:hist} gives a histogram of posterior distributions for state transition probabilities.

\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{posterior_histograms}

}

\caption{Histogram of posterior distributions for state transition probabilities.}\label{fig:hist}
\end{figure}

Figure \ref{fig:statepop} gives an example of state occupancy over time plot.
This shows that for the new algorithm there are fewer SCD and more shocks.

\begin{figure}

{\centering \includegraphics[width=20.31in]{../../images/state_pop_over_time}

}

\caption{State occupancy over time.}\label{fig:statepop}
\end{figure}

\hypertarget{cost-effectiveness}{%
\subsubsection{Cost-effectiveness}\label{cost-effectiveness}}

\begin{longtable}[]{@{}llllll@{}}
\caption{Cost-effectiveness statistics}\tabularnewline
\toprule
Strategy & Cost, \(c\) & \(\Delta c\) & QALYs, \(e\) & \(\Delta e\) & ICER \\
\midrule
\endfirsthead
\toprule
Strategy & Cost, \(c\) & \(\Delta c\) & QALYs, \(e\) & \(\Delta e\) & ICER \\
\midrule
\endhead
Baseline & & & & & \\
\textgreater{} 4\% & & & & & \\
\textgreater{} 6\% & & & & & \\
\textgreater{} 1 risk factor & & & & & \\
\textgreater{} 2 risk factors & & & & & \\
\bottomrule
\end{longtable}

Figure \ref{fig:ceac} gives the CEAC and
Figure \ref{fig:ceplane} gives the CE-plane.

\begin{figure}

{\centering \includegraphics[width=19.22in]{../../images/ceac}

}

\caption{Cost-effectiveness acceptability curve (CEAC).}\label{fig:ceac}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=19.22in]{../../images/ceplane}

}

\caption{Cost-effectiveness plane.}\label{fig:ceplane}
\end{figure}

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

Decision making as to whether have the implant is shared between patient and medical professional.
The SCD risk score is a support tool which contributes to the final decision.
An SCD risk score of 6\% is treatment recommendation for one patient versus another.
There could be individual preferences that out-weighs the risk score and mean a patient chooses the alternative.

Having more robust quality of life data will help to make such decisions.
more embedded PROMs research

Informatics Consult (Lai et al. 2021) automation to scale evidence generation and to accelerate the return of results within clinical time-scales.

We did not include CRT-D or s-ICD but that can be overcome by sensitivity analyses that show e.g.~the main message is that conclusions about the ICER etc will be robust to most departures from base assumptions other than tweaks in utility (which are the part of the model based on the weakest data!)

(Olivotto et al. 2020) The role of new DMARDs and their ability to influence quality of life.
So if both ICD and non-ICD patients experience significant increases in QoL, are these equally distributed?

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}

\begin{figure}

{\centering \includegraphics[width=0.8\linewidth]{../../images/year_of_entry_barplot}

}

\caption{Year of study entry bar plot.}\label{fig:barplot}
\end{figure}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Boriani2014}{}%
Boriani, Giuseppe, Paolo Cimaglia, Mauro Biffi, Cristian Martignani, Matteo Ziacchi, Cinzia Valzania, and Igor Diemberger. 2014. {``{Cost-effectiveness of implantable cardioverter-defibrillator in today's world}.''} \emph{Indian Heart Journal} 66 (SUPPL. 1): S101--4. \url{https://doi.org/10.1016/j.ihj.2013.12.034}.

\leavevmode\hypertarget{ref-Bryant2007}{}%
Bryant, Jackie, Hakan Brodin, Emma Loveman, and Andrew Clegg. 2007. {``{Clinical effectiveness and cost-effectiveness of implantable cardioverter defibrillators for arrhythmias: A systematic review and economic evaluation}.''} \emph{International Journal of Technology Assessment in Health Care} 23 (1): 63--70. \url{https://doi.org/10.1017/S0266462307051586}.

\leavevmode\hypertarget{ref-Caro2007}{}%
Caro, J. Jaime, Alexandra Ward, H. Baris Deniz, Judith A. O'Brien, and Jenifer L. Ehreth. 2007. {``{Cost-benefit analysis of preventing sudden cardiac deaths with an implantable cardioverter defibrillator versus amiodarone}.''} \emph{Value in Health} 10 (1): 13--22. \url{https://doi.org/10.1111/j.1524-4733.2006.00140.x}.

\leavevmode\hypertarget{ref-Colquitt2014}{}%
Colquitt, Jill L, Diana Mendes, Andrew J Clegg, Petra Harris, Keith Cooper, Joanna Picot, and Jackie Bryant. 2014. {``{Implantable cardioverter defibrillators for the treatment of arrhythmias and cardiac resynchronisation therapy for the treatment of heart failure: systematic review and economic evaluation}.''} \emph{Health Technology Assessment} 18 (56). \url{https://doi.org/10.3310/hta18560}.

\leavevmode\hypertarget{ref-Cowie2009}{}%
Cowie, Martin R., Deborah Marshall, Michael Drummond, Nicole Ferko, Michael Maschio, Matthias Ekman, Luc De Roy, et al. 2009. {``{Lifetime cost-effectiveness of prophylactic implantation of a cardioverter defibrillator in patients with reduced left ventricular systolic function: Results of markov modelling in a european population}.''} \emph{Europace} 11 (6): 716--26. \url{https://doi.org/10.1093/europace/eup068}.

\leavevmode\hypertarget{ref-Garcia-Perez2015}{}%
García-Pérez, Lidia, Pilar Pinilla-Domínguez, Antonio García-Quintana, Eduardo Caballero-Dorta, F. Javier García-García, Renata Linertová, and Iñaki Imaz-Iglesia. 2015. {``{Economic evaluations of implantable cardioverter defibrillators: a systematic review}.''} \emph{European Journal of Health Economics} 16 (8): 879--93. \url{https://doi.org/10.1007/s10198-014-0637-x}.

\leavevmode\hypertarget{ref-Lai2021}{}%
Lai, Alvina G, Wai Hoong Chang, Constantinos A Parisinos, Michail Katsoulis, M Ruth, Anoop D Shah, Vincent Nguyen, et al. 2021. {``{An Informatics Consult approach for generating clinical evidence for treatment decisions}.''} \emph{medRxiv}. https://doi.org/\url{https://doi.org/10.1101/2021.01.10.21249331}.

\leavevmode\hypertarget{ref-Magnusson2020}{}%
Magnusson, Peter, and Anders Wimo. 2020. {``{Health economic evaluation of implantable cardioverter defibrillators in hypertrophic cardiomyopathy in adults}.''} \emph{International Journal of Cardiology} 311: 46--51. \url{https://doi.org/10.1016/j.ijcard.2020.02.055}.

\leavevmode\hypertarget{ref-Mealing2016}{}%
Mealing, Stuart, Beth Woods, Neil Hawkins, Martin R. Cowie, Christopher J. Plummer, William T. Abraham, John F. Beshai, Helmut Klein, and Mark Sculpher. 2016. {``{Cost-effectiveness of implantable cardiac devices in patients with systolic heart failure}.''} \emph{Heart} 102 (21): 1742--49. \url{https://doi.org/10.1136/heartjnl-2015-308883}.

\leavevmode\hypertarget{ref-Noyes2007}{}%
Noyes, Katia, Ethan Corona, Jack Zwanziger, W. Jackson Hall, Hongwei Zhao, Hongkun Wang, Arthur J. Moss, and Andrew W. Dick. 2007. {``{Health-related quality of life consequences of implantable cardioverter defibrillators: Results from MADIT II}.''} \emph{Medical Care} 45 (5): 377--85. \url{https://doi.org/10.1097/01.mlr.0000257142.12600.c1}.

\leavevmode\hypertarget{ref-OMahony2014}{}%
O'Mahony, Constantinos O, Fatima Jichi, Menelaos Pavlou, Lorenzo Monserrat, Aristides Anastasakis, Claudio Rapezzi, Elena Biagini, et al. 2014. {``{Myocardial disease A novel clinical risk prediction model for sudden cardiac death in hypertrophic cardiomyopathy (HCM Risk-SCD)}.''} \emph{European Heart Journal} 35: 2010--20. \url{https://doi.org/10.1093/eurheartj/eht439}.

\leavevmode\hypertarget{ref-Olivotto2020}{}%
Olivotto, Iacopo, Artur Oreziak, Roberto Barriales-Villa, Theodore P. Abraham, Ahmad Masri, Pablo Garcia-Pavia, Sara Saberi, et al. 2020. {``{Mavacamten for treatment of symptomatic obstructive hypertrophic cardiomyopathy (EXPLORER-HCM): a randomised, double-blind, placebo-controlled, phase 3 trial}.''} \emph{The Lancet} 396 (10253): 759--69. \url{https://doi.org/10.1016/S0140-6736(20)31792-X}.

\leavevmode\hypertarget{ref-Ommen2020}{}%
Ommen, Steve R., Seema Mital, Michael A. Burke, Sharlene M. Day, Anita Deswal, Perry Elliott, Lauren L. Evanovich, et al. 2020. {``{2020 AHA/ACC Guideline for the Diagnosis and Treatment of Patients With Hypertrophic Cardiomyopathy}.''} \emph{Circulation} 142 (25). \url{https://doi.org/10.1161/cir.0000000000000937}.

\leavevmode\hypertarget{ref-Sanders2005}{}%
Sanders, Gillian D, Mark A Hlatky, and Douglas K Owens. 2005. {``{Cost-Effectiveness of Implantable Cardioverter--Defibrillators},''} 1471--80.

\leavevmode\hypertarget{ref-Smith2013}{}%
Smith, Tim, Luc Jordaens, Dominic A. M. J. Theuns, Pascal F. Van Dessel, Arthur A. Wilde, and M. G. Myriam Hunink. 2013. {``{The cost-effectiveness of primary prophylactic implantable defibrillator therapy in patients with ischaemic or non-ischaemic heart disease: A European analysis}.''} \emph{European Heart Journal} 34 (3): 211--19. \url{https://doi.org/10.1093/eurheartj/ehs090}.

\leavevmode\hypertarget{ref-Tomini2016}{}%
Tomini, F. 2016. {``{A review of economic evaluation models for cardiac resynchronization therapy with implantable cardioverter defibrillators in patients with heart failure}.''} \emph{Eur J Health Econ}, 1159--72. \url{https://doi.org/10.1007/s10198-015-0752-3}.

\leavevmode\hypertarget{ref-Yao2007}{}%
Yao, Guiqing, Nick Freemantle, Melanie J Calvert, Stirling Bryan, Jean-claude Daubert, and John G F Cleland. 2007. {``{The long-term cost-effectiveness of cardiac resynchronization therapy with or without an implantable cardioverter-defibrillator}.''} \emph{European Heart Journal}, 42--51. \url{https://doi.org/10.1093/eurheartj/ehl382}.

\leavevmode\hypertarget{ref-Zhang2014}{}%
Zhang, Zhiyong. 2014. {``{WebBUGS: Conducting Bayesian statistical analysis online}.''} \emph{Journal of Statistical Software} 61 (7): 1--30. \url{https://doi.org/10.18637/jss.v061.i07}.

\leavevmode\hypertarget{ref-Zomer2017}{}%
Zomer, Ella, David Osborn, Irwin Nazareth, Ruth Blackburn, Alexandra Burton, Sarah Hardoon, Richard Ian, et al. 2017. {``{Effectiveness and cost-effectiveness of a cardiovascular risk prediction algorithm for people with severe mental illness ( PRIMROSE )}.''} \emph{BMJ Open}, 1--11. \url{https://doi.org/10.1136/bmjopen-2017-018181}.

\end{CSLReferences}

\end{document}

### 2021-07-21 HCM.SCD.CEanalysis/R/annual_trans_counts.R

@@ -1,10 +1,14 @@
 
-#' annual_trans_counts
+#' Annual transition counts
 #'
 #' Aggregate individual data to annual totals.
 #'
-#' @param data
-#' @param cycle_length
+#' @param data Individual patient data with
+#'              cvs_all:
+#'              scd:
+#'              non_cvs:
+#'              time:
+#' @param cycle_length Break points length
 #' @export
 #'
 annual_trans_counts <- function(data,
@@ -18,10 +22,12 @@ annual_trans_counts <- function(data,
       non_scd = as.numeric((cvs_all & !scd) | non_cvs),
       cens = as.numeric(!non_scd & !scd)) %>%
     select(time, non_scd, scd, cens) %>%
+    # year of event
     mutate(yr_grp = cut(time,
                         breaks = seq(0, 40, by = cycle_length),
                         right = FALSE)) %>%
     group_by(yr_grp) %>%
+    # sum totals by year
     summarise(scd_count = sum(scd),
               non_scd_count = sum(non_scd),
               cens_count = sum(cens)) %>%
@@ -31,6 +37,7 @@ annual_trans_counts <- function(data,
            healthy = n_pop - cum_cens - cum_scd - cum_non_scd,
            at_risk0 = lag(healthy, default = n_pop),
            # actuarial assumption
+           # half censored in at risk group
            at_risk = round(at_risk0 - 0.5*cens_count, 0))
 }
 

### 2021-07-21 HCM.SCD.CEanalysis/R/obs_aggr_trans_mat.R

@@ -1,12 +1,14 @@
 
-#' obs_aggr_trans_mat
+#' Aggregate transition matrix
 #'
 #' Create aggregate observed transition counts
 #' see p.185 Table 5.2 chapter 5 in BCEA book
 #'
-#' Assume 3 state model: "healthy", "scd_count", "non_scd_count"
+#' Assume 3 state model:
+#' 1) healthy 2) scd_count 3) non_scd_count
 #'
 #' @param data new state arrivals at time step
+#'             result of running `annual_trans_counts()`
 #'             e.g. data = data_rule$`FALSE`
 #' @param max_year
 #' @export
@@ -30,17 +32,17 @@ obs_aggr_trans_mat <- function(data,
     trans_mat[[i + 1]] <- trans_mat[[i]]
 
     # new transitions
-      trans_mat[[i + 1]]["healthy", state_names] <-
-        trans_mat[[i + 1]]["healthy", state_names] +
-        unlist(data[i, state_names])
+    trans_mat[[i + 1]]["healthy", state_names] <-
+      trans_mat[[i + 1]]["healthy", state_names] +
+      unlist(data[i, state_names])
 
-      trans_mat[[i + 1]]["scd_count", "scd_count"] <-
+    trans_mat[[i + 1]]["scd_count", "scd_count"] <-
       trans_mat[[i + 1]]["scd_count", "scd_count"] +
-        trans_mat[[i]]["healthy", "scd_count"]
+      trans_mat[[i]]["healthy", "scd_count"]
 
-      trans_mat[[i + 1]]["non_scd_count", "non_scd_count"] <-
+    trans_mat[[i + 1]]["non_scd_count", "non_scd_count"] <-
       trans_mat[[i + 1]]["non_scd_count", "non_scd_count"] +
-        trans_mat[[i]]["healthy", "non_scd_count"]
+      trans_mat[[i]]["healthy", "non_scd_count"]
   }
 
   names(trans_mat) <- c(0:max_year)

### 2021-07-21 HCM.SCD.CEanalysis/scripts/main_ce_analysis.R

@@ -27,7 +27,6 @@ data("ipd_risk")
 # (1.icd  2.shock 3.mortality)
 # (4.~icd 5.scd   6.mortality)
 
-
 # start state populations
 init_risk6 <- table(ipd_risk$risk_over_6)
 init_obs <- table(ipd_risk$icd)
@@ -40,33 +39,35 @@ n_init$risk6 <- c(init_risk6["TRUE"], 0, 0,
 
 n_sim <- nrow(lambda.0)
 S <- length(n_init$obs)
-J <- 12  # max time (year)
+J <- 12                 # max time (year)
 
 
 # utilities and costs --------
 # for each intervention
 
-u_hcm <- 0.637   # managed with an ICD (Noyes 2007)
-u_icd <- -0.05   # ICD annual detriment
-u_shock <- 0.5   # annual detriment
-u_icd_init <- -6/365 # icd implant one-off (Smith EJH 2013)
+u_hcm <- 0.637       # managed with an ICD (Noyes 2007)
+u_icd <- -0.05       # ICD annual detriment
+u_shock <- 0.5       # annual detriment
+u_icd_init <- -6/365 # ICD implant one-off (Smith EJH 2013)
 
-e_unit <- list(obs =   c(u_hcm + u_icd, u_shock, 0,
-                         u_hcm,         0,       0),
-               risk6 = c(u_hcm + u_icd, u_shock, 0,
-                         u_hcm,         0,       0))
+e_unit <-
+  list(obs =   c(u_hcm + u_icd, u_shock, 0,
+                 u_hcm,         0,       0),
+       risk6 = c(u_hcm + u_icd, u_shock, 0,
+                 u_hcm,         0,       0))
 
 
-# non-fatal HCM related events (UK Stroke Association)
-c_nfatal <- 22880
+c_shock <- 22880      # non-fatal HCM related events (UK Stroke Association)
 c_icd <- 4666         # EY02B NHS tariffs implantation cost
-c_rscore <- 20
-c_icd_appt <- 10
-c_icd_repl <- 45000
+c_rscore <- 20        # cost of admin risk score
+c_icd_appt <- 10      # cost ICD initial appointment
+c_icd_repl <- 45000   # cost ICD replacement appointment
 
 # implant complication
-p_compl <- 0.047        # Smith EHJ 2013
-p_compl_repl <- 0.032   # Smith EHJ 2013
+# Smith EHJ (2013)
[…]

### 2021-07-21 HCM.SCD.CEanalysis/scripts/prep_study_data.R

@@ -29,15 +29,16 @@ data("data_set1")
 CYCLE <- 1 #year
 
 # non-deterministic decision rule
-FUZZY_RISK <- TRUE
+# random at boundary
+FUZZY_RISK <- FALSE
 
 fuzzy_noise <-
   if (FUZZY_RISK) {
     rnorm(nrow(data_set1), 0, 0.001)
   } else {0}
 
-
 # subset cohort
+# by number of risk factors
 rf_threshold <- 1
 
 ipd_risk <-

### 2021-07-21 HCM.SCD.CEanalysis/inst/rmarkdown/draft.Rmd
check that probabilities look sensible and written ggplot version of trans probs histograms

@@
 
-Figure \@ref(fig:hist) gives a histogram of posterior distributions for state transition probabilities.
+Figures \@ref(fig:histnoICD) and \@ref(fig:histICD) give histograms of posterior distributions for state transition probabilities for no groups without and with ICD implants.
 
-```{r hist, fig.align="center", echo=FALSE, out.width = "80%", fig.cap="Histogram of posterior distributions for state transition probabilities."}
-knitr::include_graphics("posterior_histograms.png")
+```{r histnoICD, fig.align="center", echo=FALSE, out.width = "80%", fig.cap="Histogram of posterior distributions for state transition probabilities."}
+knitr::include_graphics("../../images/post_hist_noICD.png")
+```
+
+```{r histICD, fig.align="center", echo=FALSE, out.width = "80%", fig.cap="Histogram of posterior distributions for state transition probabilities."}
+knitr::include_graphics("../../images/post_hist_withICD.png")
 ```

### 2021-07-21 HCM.SCD.CEanalysis/inst/rmarkdown/draft.tex
check that probabilities look sensible and written ggplot version of trans probs histograms

@@
 
-Figure \ref{fig:hist} gives a histogram of posterior distributions for state transition probabilities.
+Figures \ref{fig:histnoICD} and \ref{fig:histICD} give histograms of posterior distributions for state transition probabilities for no groups without and with ICD implants.
 
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{posterior_histograms}
+{\centering \includegraphics[width=0.8\linewidth]{../../images/post_hist_noICD}
 
@@
 
-\caption{Histogram of posterior distributions for state transition probabilities.}\label{fig:hist}
+\caption{Histogram of posterior distributions for state transition probabilities.}\label{fig:histnoICD}
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.8\linewidth]{../../images/post_hist_withICD}
+
+}
+
+\caption{Histogram of posterior distributions for state transition probabilities.}\label{fig:histICD}
 \end{figure}

### 2021-07-21 HCM.SCD.CEanalysis/BUGS/script.R

@@ -14,6 +14,9 @@
 # 5 scd      0 0 0 0 1 0
 # 6 death    0 0 0 0 0 1
 
+# lambda.0: observed
+# lambda.1: > 6% risk group
+
 
 library(R2jags)
 

### 2021-07-21 HCM.SCD.CEanalysis/scripts/main_ce_analysis.R

@@ -42,7 +42,7 @@ S <- length(n_init$obs)
 J <- 12                 # max time (year)
 
 
-# utilities and costs --------
+#### utilities
 # for each intervention
 
 u_hcm <- 0.637       # managed with an ICD (Noyes 2007)
@@ -56,17 +56,13 @@ e_unit <-
        risk6 = c(u_hcm + u_icd, u_shock, 0,
                  u_hcm,         0,       0))
 
+### costs
 
 c_shock <- 22880      # non-fatal HCM related events (UK Stroke Association)
 c_icd <- 4666         # EY02B NHS tariffs implantation cost
-c_rscore <- 20        # cost of admin risk score
+c_rscore <- 0#20        # cost of admin risk score
 c_icd_appt <- 10      # cost ICD initial appointment
 c_icd_repl <- 45000   # cost ICD replacement appointment
-
-# implant complication
-# Smith EHJ (2013)
-p_compl <- 0.047        # initial
-p_compl_repl <- 0.032   # replacement
 
 # weighted infection, dislodgement
 c_compl <- (0.0227*37116 + 0.00828*6146)/(0.0227 + 0.00828)
@@ -75,6 +71,11 @@ c_entry <- list(obs =   c(0, 0, 0,
                           0, 0, 0),
                 risk6 = c(0, 0, 0,
                           0, 0, 0))
+
+# implant complication
+# Smith EHJ (2013)
+p_compl <- 0.047        # initial
+p_compl_repl <- 0.032   # replacement
 
 # time to replace ICD
 t_repl <- 10
@@ -155,6 +156,8 @@ res_new <-
          e_unit,
          rc_init(c_icd, c_rscore),
          pdecr)
+
+save(res_new, file = "data/res_new.RData")
 
 # total cost across all time points
 c <-

### 2021-07-21 HCM.SCD.CEanalysis/scripts/pop_counts_plot.R

@@ -1,5 +1,6 @@
 
-# pop counts plot
+# state pop counts plot
+# over time
 # stacked percentage bar plot
 
 
@@ -65,6 +66,8 @@ facet_plot <-
   geom_bar(position = "fill", stat = "identity") +
   scale_x_continuous(breaks = 1:12) +
   facet_grid(.~strat)
+
+facet_plot
 
 ggsave(facet_plot, file = "images/state_pop_over_time.png")
 

### 2021-07-21 HCM.SCD.CEanalysis/scripts/posterior_hist_plots.R

@@ -1,30 +1,95 @@
 
 # posterior histograms
 # transition probabilities
-#
 
 
 data("jags_output")
-R2WinBUGS::attach.bugs(mm1$BUGSoutput)
+R2jags::attach.jags(mm1)
 
+# with ICD
 x11()
 par(mfrow = c(2,3))
+# obs
 hist(lambda.0[, 1, 1], breaks = 20, xlab = "Prob", main = "ICD -> ICD")
 abline(v = mean(lambda.0[, 1, 1]), col = "red")
 hist(lambda.0[, 1, 2], breaks = 20, xlab = "Prob", main = "ICD -> Shock")
 abline(v = mean(lambda.0[, 1, 2]), col = "red")
 hist(lambda.0[, 1, 3], breaks = 20, xlab = "Prob", main = "ICD -> All-cause death")
 abline(v = mean(lambda.0[, 1, 3]), col = "red")
+# > 6%
+hist(lambda.1[, 1, 1], breaks = 20, xlab = "Prob", main = "ICD -> ICD")
+abline(v = mean(lambda.1[, 1, 1]), col = "red")
+hist(lambda.1[, 1, 2], breaks = 20, xlab = "Prob", main = "ICD -> Shock")
+abline(v = mean(lambda.1[, 1, 2]), col = "red")
+hist(lambda.1[, 1, 3], breaks = 20, xlab = "Prob", main = "ICD -> All-cause death")
+abline(v = mean(lambda.1[, 1, 3]), col = "red")
 
+# low risk
+x11()
+# obs
+par(mfrow = c(2,3))
 hist(lambda.0[, 4, 4], breaks = 20, xlab = "Prob", main = "Low risk -> Low risk")
 abline(v = mean(lambda.0[, 4, 4]), col = "red")
 hist(lambda.0[, 4, 5], breaks = 20, xlab = "Prob", main = "Low risk -> SCD")
 abline(v = mean(lambda.0[, 4, 5]), col = "red")
 hist(lambda.0[, 4, 6], breaks = 20, xlab = "Prob", main = "Low risk -> All-cause death")
 abline(v = mean(lambda.0[, 4, 6]), col = "red")
+# > 6%
+hist(lambda.1[, 4, 4], breaks = 20, xlab = "Prob", main = "Low risk -> Low risk")
+abline(v = mean(lambda.1[, 4, 4]), col = "red")
+hist(lambda.1[, 4, 5], breaks = 20, xlab = "Prob", main = "Low risk -> SCD")
[…]
